### 2017-12-27 test.Rmd
stash

---
title: "test"
author: "Ross"
date: "19/12/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\documentclass[12pt, english, oneside]{amsart}
\begin{document}

Figure \ref{fig:plot} doesn't have it's label.

<<plot>>=
plot(x=0, y=0)
@

Figure \ref{fig:plot2} has its label.

\begin{figure}
\caption{\label{fig:plot2}}
<<>>=
plot(x=1,y=1)
@
\end{figure}

\end{document}