### 2018-09-17 Second.Rnw
big commit, getting close!

\documentclass{article}

\usepackage[backend=bibtex, sorting=none, citestyle=numeric]{biblatex}
\usepackage[margin=0.8in]{geometry}
\bibliography{references}

%  For caption width
\usepackage{graphicx,caption}
% For Centering figs
\usepackage[export]{adjustbox}
% For split figures
\usepackage{subcaption}
%  appendix
\usepackage[titletoc,page]{appendix}
\setlength{\parindent}{0em}

\usepackage{tikz,lipsum,lmodern}
\usepackage[most]{tcolorbox}

% for cover page
\usepackage{geometry}
\usepackage{xcolor}
\usepackage{graphicx}

% \usepackage{draftwatermark}
% \SetWatermarkText{DRAFT}
% \SetWatermarkScale{1.5}
% \SetWatermarkLightness{0.95}

\title{Efficiency}
\author{Aidan Morrison}

% \SweaveOpts{concordance=TRUE}
\definecolor{titlepagecolor}{cmyk}{0, 0, 0, .30}
\definecolor{namecolor}{cmyk}{0, 0, 0, .80}

\begin{document}

%-----------------------------------------------------------------

% ----------------------------------------------------------------
\begin{titlepage}
\newgeometry{left=7.5cm} %defines the geometry for the titlepage
\pagecolor{titlepagecolor}
\noindent

{\huge \textsf{A response to the Government response}}
\vskip\baselineskip
{\huge \textsf{(Or lack thereof)}}
\vskip\baselineskip
{\huge \textsf{on the question of the pumpjet}}
\vskip\baselineskip
\noindent
\textbf{Aidan Morrison}
\vskip\baselineskip
\textsf{August 2018}
\vskip\baselineskip
\vskip\baselineskip

\vfill
\noindent

\includegraphics[width=6cm]{tl_logo.png}\\[-1em]
\color{white}
\makebox[0pt][l]{\rule{1.3\textwidth}{1pt}}
\par
\noindent
\textbf{\textsf{Trendlock}} \textcolor{namecolor}{\textsf{Consulting}}

\end{titlepage}
\restoregeometry % restores the geometry
\nopagecolor% Use this to restore the color pages to white

\newpage

<<setup, include=FALSE, cache=FALSE, echo=FALSE>>=
opts_chunk$set(fig.path='figures/plots-', fig.align='center', fig.show='hold', eval=T, echo=F)
options(replace.assign=TRUE,width=80)
library(tidyverse)
library(submarines)
@

\newpage

\tableofcontents

\newpage
\setlength{\parskip}{1em}

\section{Introduction} \label{intro}

The pumpjet propulsion system is a key feature of Naval Group's Shortfin Barracuda submarine concept, and was apparently a key reason why the concept was selected by the Australian Government as the basis for the design of Australia's new submarine under project SEA 1000.  Naval Group's official website, reinforced by comments from officials representing both Naval Group and the Australian Government,  claims that the pumpjet system offers acoustic advantages which will make the Shortfin Barracuda much more effective than a submarine driven by a conventional propeller.

However, over a number of years now the pumpjet choice has attracted sustained attention and criticism, mostly around its efficiency at low speeds, where a diesel-electric submarine would necessarily spend a great deal of its operating time.  I was commissioned by Gary Johnston of Submarines for Australia to investigate this topic in detail and produce a report \parencite{morrison2018}, which concluded that a very substantial reduction in dived range and endurance (as well as overall range) would necessarily result from the pumpjet choice.  Furthermore, that paper challenged the assumption that the pumpjet would be acoustically superior at low speeds, where open propellers would have no cavitation, and cause much less turbulence than the equivalent jet.

That paper, and a subsequent blog post \parencite{morrison2018a} on the topic which made the paper publically available, attracted the interest of the Senate Economics Reference Committee Inquiry into the Future of Naval Shipbuilding in Australia.  Senator Chris Ketter asked written questions on notice to the Department of Defence requesting responses to the paper specifically, and related comments from public officials, with a return date of 3rd April 2018.  I was also invited to provide evidence to the committee in a public hearing on the 7th of June 2018.  In that public hearing, I stood by the conclusion of my substantive paper, which was: ``between two otherwise identical submarines, the one with the pump jet will always have a lower dived endurance, a lower dived range, a worse indiscretion ratio, a lower overall endurance and a lower overall range than the one with the propeller."  In the presence of both the outgoing acting CEO of Naval Group Australia Brent Clark, and General Manager of Submarines (for Defence) Stephen Johnson, I pointed out that that statement was yet to be challenged, and that I'd be very surprised if anyone with technical competence or professional integrity sought to challenge it.

In their own evidence during the course of the day, neither of those two did choose to challenge it.  Senator Kim Carr requested that defence produce a written response to that report for the committee prior to their handing down of their finaly report for the Inquiry, and Defence agreed to do their best in the time available, with the head of Capability and Sustainment Group (CASG) Kim Gillis noting that ``we do hold that expertise", and Senator Carr emphasising that ``it would look pretty bloody stupid after the report's handed in''. Despite these exhortations for timeliness, in the final report of the committee, this request, and defence's acceptance of it was noted, but no acknowledgement was made of whether any response was actually received, or what it contained \parencite{SERC2018final}.

In the August edition of Australian Defence Magazine a feature article on the titled `Future Submarine: Pumpjet propulsor technology choice still not well explained' \parencite{kerr2018} appeared, referencing both my testimony in the hearing and the paper.  Comment for the article was sought from CASG, the Royal Australian Navy (RAN) and Naval Group, and none was forthcoming.  However, some substantial comments were provided by Dr David Kershaw, the chief of DSTG's Maritime Division, who was also Science and Technology advisor for a time on SEA 1000, the Future Submarine Project.

This paper has been commissioned by Gary Johnston to provide a summary of the responses that have been received from official sources (in government, and Naval Group) to the earlier major report, and respond to them.  At the time of its original commissioning, two substantive arguments emerge seeking to engage with the technical arguments made in the earlier paper.  The first is that the large size of the submarine now makes the pumpjet viable, where as for smaller submarines the large weight of the pumpjet would make it unviable, and that the pumpjet would be generally well-suited to smaller sumbarines as well were it not for this weight consideration.  The second is that interaction between the pumpjet and the hull obviate the evidence I have produced showing that pumpjet efficiency declines at low speeds, which are claimed to neglect considering such interactions.

In early September, the Department of Defence gave a six-page written response to my paper, a full three months after the initial request. The two substantive arguments already identified were repeated, but the weight of their argument actually came to rest on another assumption.  This assumption was mentioned already by Dr Kershaw in the ADM article, but it was so completely at odds with the most basic physics that I was inclined to dismiss it as perhaps a clumsily worded statement made in an interview compounded by a little journalistic mis-quotation, rather than a substantive argument.   The re-appearance of this argument, which was spelled out and repeated multiple times in a written document that was prepared by Defence over the course of three months, well after any nominal deadline for timeliness had elapsed, dispelled all illusions that the argument was included by accident.  As such, I have also addressed it here as the first of three substantial arguments made against my earlier conclusions regarding pumpjet efficiency, which has quite significantly altered the conclusions of this paper, which seeks to consider all three lines of reasoning in sufficient detail to suitably discuss the circumstances in which such arguments might be made.

This paper commences with a chronologically sequenced history of the most significant comments that have been made by public officials or Naval Group/DCNS with respect to the pumpjet, highlighting the most important implications of each in their context.  These are followed by three sections, addressing in more detail the three substantive arguments against my conclusions.  Some discussion developing the final conclusions follows, followed by appendices which include more substantial extracts or originals sources where appropriate, for reference.

\section{A brief chronology of comment}

The following gives a brief overview of the significant known public comments regarding the pumpjet that have been made by official representatives of either the Commonwealth of Australia, or Naval Group, formally known as DCNS.  Only brief quotations are selected here, with references being made to the full excerpts either in Appendices or References.  Any emphasis in-text has been added.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=red!5!white,colframe=red!75!black, title= \textbf{28 May 2012 - Senate Estimates}]
\textit{Senator Johnston directly challenges Rear Admiral Justin Moffit (then director of SEA 1000) about the pumpjet proposal.  Full exchange in \ref{SENMoffit}. Emphasis added.}
\\
\\
\textbf{Senator JOHNSTON:} I am told that there are no pump jets on any conventional class submarines in the world.
\\
\textbf{Rear Adm. Moffitt:} That is quite right. It might be a very easy task for us to write it off. I do not know, because I have not done-
\\
\textbf{Senator JOHNSTON:} The French and the Russians have tried. No-one has used them. They are not being used on anything other than nuclear powered submarines-particularly the Americans.
\\
\textbf{Rear Adm. Moffitt:} \textbf{If it does not look like it makes sense, I will rule it out. But I cannot give you the answer to that yet because we have not done the work.}
\\
\textbf{Senator JOHNSTON:} Who had the idea to look at pump jets? Where did that come from? Did it come from capability group saying that we had better had a look at that?
\\
\textbf{Rear Adm. Moffitt:} I do not know the answer to that. I doubt that there is a clear answer.

\end{tcolorbox}

This encounter shows that the viability of pumpjets was challenged some time ago.  It is noteworthy at the time that RADM Moffit makes no attempt to defend the consideration, acknowledging even that it might be easy for them to ``write it off".  It is also noteworthy the tone of confidence that Senator Johnston's challenge has, driving directly for an explanation of who had the original idea, which RADM Moffit was unable or unwilling to answer.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{10 Feb 2016 - Sean Costello, CEO DCNS Australia}]
\textit{In a blog post\parencite{costello2016} for the ASPI Strategist prior to the announcement in April, Sean Costello describes the significance of the pump jet technology to their bid. Emphasis added.}
\\
\\
``As a complete submarine power, we understand conventional propulsion, which is why we also understand that propulsion is but one part of the submarine puzzle. In designing to the Australian requirement, it should come as no surprise then that \textbf{the conventionally powered Shortfin Barracuda Block 1A is only 5\% lighter than its nuclear cousin.}"
\\
\\
``There's no better example of what I'm talking about than the \textbf{offer from France to transfer to Australia sovereign control and use of pump-jet propulsion technology for the Shortfin Barracuda}-technology resident only in France, the UK and the USA. Technology born from the French SSBN program a generation ago. The stealth and hydrodynamic performances of pump jet propulsion are of course classified and in Australia known only to DCNS and the Australian Government."

\end{tcolorbox}

The significance of the pumpjet in Frances bid was clearly acknowledged well before the public announcement of DCNS's success.  It also provides a clear indication that the size of the submarine already appeared to be quite well-known, and was only intended to be a very minor modification from the nuclear Barracuda.  For such a small variation in displacement, it might be very reasonably expected that the length of the submarine would vary (the easiest modification), but the diameter would not need to.  These comments also display quite clearly that the pumpjet technology in France's possession is recognised as common with technology in the US and the UK.  This rules out the likelihood that some special or unique pumpjet technology, better suited just to diesel-electric submarines, had been pioneered by France alone.  This comment also directly implies that information regarding the performance of the pumpjet had already been given to the Australian  government, \textit{prior} to the conclusion of the Competitive Evaluation Process (CEP).

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{8 Apr 2016 - Sean Costello, CEO DCNS Australia, and Gerard Autret, Chief Naval Architect of the Shortfin Barracuda}]
\textit{In a blog post\parencite{autret2016} for the ASPI Strategist prior to the announcement in April, Sean Costello describes a number of the major features and design approaches for the Shortfin Barracuda, including relating the choice of the pumpjet closely to an accoustic requirement to close with another submarine at higher speed.}
\\
\\
``The challenge for any attack submarine is to maintain a nearly silent acoustic signature at a speed necessary to manoeuvre within weapons range of the target. The nuclear Barracuda is designed to \textbf{reduce radiated noise when operating at a speed sufficient in order to close a threat submarine undetected. Of particular importance is the pump-jet propulsor}, which combines a rotor and stator within a duct to significantly reduce the level of radiated noise through the effects of wake harmonisation and avoidance of cavitation."
\end{tcolorbox}

This comment provides clear context for the importance of the pump-jet.  Consistent with the physics outlined in Reference \parencite{morrison2018} the acoustic advantage is only emphasised by DCNS as being significant at higher speeds.  Whether it is viable for a diesel-electric submarine (as opposed to the nuclear one mentioned here) to maintain such speeds to close with many types of targets is left unclear, or at least unmentioned.  As later discussed in my evidence to the Senate \ref{SENmorrison}, this is consistent with DCNS being well aware with the advantages and disadvantages of the pumpjet, whilst not making verifyably false claims.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{May 2016 onwards - DCNS/Naval Group Website for Shortfin Barracuda}]
\textit{The official website for the submarine claims the advantages of the pumpjet in specific terms, drawing comparisons with ``obsolete propeller technology" and ``otherwise identical submarines".  Last accessed 4 Sept 2018 \parencite{naval2018}. Emphasis added.}
\\
\\
``Pump jet propulsion means the Shortfin Barracuda can move more quietly than submarines with obsolete propeller technology. In a confrontation between two \textbf{otherwise identical submarines}, the one with pump jet propulsion always has the \textbf{tactical} advantage."

\end{tcolorbox}

The official explanation of the designer and builder's website makes a carefully worded explanation of the superiority of the pumpjet, by limiting the comparison to a ``tactical" encounter with an ``otherwise identical" submarine.  As discussed in \ref{SENmorrison}, this is the only \textit{possible} target that an elevated speed - such as could be achieved by a diesel-electric submarine - might assist in closing with.  The submarine would be at a substantial `strategic' disadvantage in such an encounter, since its lower dived endurance, dived range, and poorer indiscretion ratio dramatically increases the probability that the pumpjet propelled submarine would wind up being the hunted, rather than the hunter.  More significantly, the speeds achieved would not allow the submarine to close with any other target from behind such as almost all surface ships, or nuclear submarines.  This statement exhibits extreme care in the wording in order to be plausibly defensible, and is aguably misleading.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{4 October 2017 - Jean-Michel Billig, Executive Director Australian submarine Program, Naval Group}]
\textit{In comments to journalists at the PACIFIC 2017 trade show, the Naval Group executive in charge of the Australian Program acknowledges that convetional propellers may be used on the Shortfin Barracuda \parencite{sbs2017}.  Subsequent reports quote Naval Group shifting responsibility for the propulsion choice to the Australian Government.}
\\
\\
(Interview transcribed from Reference \parencite{lane2017}.)

\textbf{Journalist:} Does the pumpjet technology that is on Suffren remain viable for the Australian submarine?

\textbf{Jean-Michel Billig:} It could be viable...

\textbf{Journalist:} So it's not mandated, its...

\textbf{Jean-Michel Billig:} What is driving the choice of technology is basically the operational requirement...

\textbf{Journalist:} So we could end up with conventional propeller?

\textbf{Jean-Michel Billig:} Yes, yes, we could, we could, we could also end up with something different
\\
\\
``However decisions on forms of propulsion are ultimately a matter for the Australian Government."\parencite{iggulden2017}
\end{tcolorbox}

This crucial concession reveals a very different level of confidence in the pumpjet being suitable for the Australian diesel-electric submarine.  It also gives an important glimpse into the processes that selected the winning bidder, and key technology components, which reveals a clear contradictions between what would have been previously foregone conclusions.  By revealing that the pumpjet has not yet been firmly decided upon, it is directly implied that France's exclusive offer to transfer the pumpjet technology could not have been the real distinguishing feature that elevated that bid above the others.  Simultaneously, the statement that technology choice is a product of the operational requirement makes it clearly implies that the operational requirement has not yet been determined with sufficient precision to make the suitable technology.  As a ship-builder with ample experience in both technologies, uniquely positioned as having fitted both pumpjets and propellers to both diesel and nuclear submarines, it is remakeable that the performance information transferred prior or during the CEP was not sufficient to allow a decision  on the technology to be made.  Moreover, it also reveals that this was not only the case at the conlcusion of the CEP, and remained so well over a year after.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{25 Oct 2017 - Hevre Guillou, CEO Naval Group}]
\textit{In an interview that was published in a post\parencite{guillou2017} on the ASPI Strategist the global CEO of Naval group responds to a question from Brendan Nicholson asking whether the Australian submarine will be fitted with a pump jet, following Jean-Michel Billig's comments. Emphasis added.}
\\
\\
``Naval Group is the manufacturer of both conventional propeller-driven submarines and larger nuclear submarines with the propulsor. The propulsor's inclusion means the future submarine will move more quietly than submarines with obsolete propeller technology. \textbf{In a contest between two otherwise identical submarines, the one with the pump jet always has the tactical advantage. The propulsor was an integral part of our bid. The Department of Defence says it intends to use the pump jet propulsion system.}"
\end{tcolorbox}

The verbatim quotation of the carefully worded comparison between ``otherwise identical" submarines by the global CEO more than a year after they first appeared on the Australian DCNS website demonstrate a high degree of preparation, and continued careful word selection on the topic of the pumpjet.  This statement also further reveals nuances about the CEP selection process, simultaneously declaring that the propulsor (i.e. pumpjet) was ``integral" to their bid, whilst also deferring to what the Department of Defence ``intends" do use. This further exposes a clear tension between how 'integral' the pumpjet was to the bid, and Defence's freedom to choose otherwise.  It also exposes a dramatic divergence between the level of emphasis that is given to the pumpjet as an important key feature of the proposal in public promotion (as highlighted in \cite{lane2017}) and a Defence decision which seemed to choose only to have the option of including the pumpjet, with the real decision to be deferred.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=red!5!white,colframe=red!75!black, title= \textbf{25 Oct 2017 - Senate Estimates}]
\textit{Senator Hanson directly challenges Rear Admiral Gregory Sammut (then director of SEA 1000) about the pumpjet proposal.  Full exchange in \ref{SENhanson}.}
\\
\\
\textbf{Senator Hanson:}   This is why you've chosen them as the design-to go with the pump
jet. So it's not propellers, is it?

\textbf{Rear Adm. Sammut:}   The proposal offered by France included a pump jet, and that's the
proposal that we've accepted. \textbf{A pump jet remains on offer, and that's what we intend to use
for the Future Submarine.}

\textbf{Senator Hanson:}   And that's why you chose France's Naval Group to-

\textbf{Rear Adm. Sammut:}   No, that is not, alone, why we chose France. There are a number of
things. I am on record as providing testimony at the Senate of the number of things we
considered in the selection of Naval Group as our international partner. There were a range of
factors, and they were all over the capability of the international partner to deliver a regionally
superior submarine and one which Australia would have the sovereign capacity to operate and
sustain throughout its life. There were a number of criteria in mind.
We looked at the capabilities. That includes the platform capabilities. We looked at the
ability of the international partners to work with the combat system integrator of our choice.
Those things in particular pertain to the capability of the boat. We looked at rough order of
magnitude costs and schedules to ensure the participants understood the scale of the program
that we were embarking upon and could give us insights into the manner in which they would
manage costs and schedules throughout the program. We looked at matters concerning
program management. \textbf{We looked at the design and the way that safety is incorporated into
the design. It's important that that was understood from the outset because, whilst the
international partner would be designing the submarine, it would be the Commonwealth that
would be responsible for certifying the submarine as safe and fit for operations, and we
needed to make sure that the certifying process would match the design philosophy.}
We looked at methods of sustainment. \textbf{We looked at the manner in which each of the
participants would support the establishment of a sovereign sustainment capability in
Australia.} Of course, that was a key lesson from the Collins program. Even though we built
the submarines in Australia, it took us some time to establish the sovereign capacity to sustain
those boats. Indeed, having done that now is the reason why we are able to deliver beyond
international benchmark availability. \textbf{We also looked into the way in which data would be
transferred to Australia, the manner in which intellectual property would also be managed in
the program.} We looked at the ability of the participants to manage risks. Again, throughout
this complex program, that would be a very important factor. I would say that we didn't just
pick a submarine with a pump jet; we picked an international partner that could meet a range
of requirements to achieve not only a regionally superior submarine but the sovereign
capacity to operate and sustain that submarine and to deliver that within the time frame that
would be necessary.
\\
\\
...
\\
\\
\textbf{Rear Adm. Sammut:}  ...  The effort is not to convert an
existing boat into a diesel-electric submarine; we're designing a submarine using many of the
learnings and references that come out of the Barracuda design. \textbf{A good example there would
be something like hull diameter. If you're going to change the diameter of the hull of a
submarine, you're into a whole new set of calculations about the structural integrity and
strength of the submarine.} If we can utilise the work that France has already done in having a
hull diameter that's very similar to Barracuda, \textbf{we can leverage all of that design work, which
has not only gone into the Barracuda but also has been tested by France in their methods of
design and certification to get to this point.}

\end{tcolorbox}
This dialogue in the Senate Estimates received media coverage highlighting Senator Hanson's apparent lack of understanding about the difference between a submarine's power source (nuclear or diesel) and propulsion (propeller or pumpjet).  However, she succeeded in applying sustained pressure to the question of the pumpjet, and the significance that it had in the selection of DCNS as a design partner.  This elicited a committment to the pumpjet that appeared hesitant in its wording, which noted that the pump jet ``remains on offer", which is what ``we intend to use".

RADM Sammut goes to length to explain that the pumpjet was not ``alone" the reason for selecting DCNS as a partner, and in doing so reveals important insights into Defence's thinking and process regarding the decision.  Crucially, he reveals that the `diameter of the hull' is an important aspect of France's design reference in the Barracuda, pointing out that changing the diameter of the hull incurrs a ``whole new set of calculations about the structural integrity and strength of the hull."  Together with remarks about the ``way that safety is incorporated into the design" and concerns about the alignment of the certifying process with the ``design philosophy" it would appear that the hull diameter actually emerges as plausibly one of the most significant distinguishing features in the French bid, perhaps alongside the pumpjet, and aligns with emphasis given to the exact diameter of the pumpjet remaining unchanged in media report earlier in the year.\parencite{kerr2017}

This is alarming for two reasons.  The first is that it contradicts subsequent explanations regarding pumpjet efficiency being incorrectly assessed by outsiders because they failed to account for how how the hull would be `tuned' to the pumpjet and the pumpjet `tuned' to the hull.  If the hull diameter is already fixed and known as being identical to that of the nuclear Barracuda, then no adjustment ought to be necessary to a good working pumpjet design that is used to propel this platform on the hull's account.  The performance of the pumpjet on the relevant hull would be well understood by the manufacturer offering it, and presumeably available to the Australian government who were looking to acquire it.  Further discussion of the hull-propulsor interaction is given in Section \ref{hullpropulsor}.

The second is that is reveals a stunning naivety in terms of basic understanding of the heirachy of engineering challenges.  It may be true that there are very many complex and challenging engineering calculations involved in building a submarine. However, it is certainly not true that scaling up a pressure hull to a new diameter to meet the same fixed static load is amongst the most difficult, risky or complex.  Whilst extending the length of a submarine is undoubtedly even easier, adjusting the diameter still to meet a structural integrity requirement to meet a known static pressure (driven by the dive depth) is a relatively simple calculation in comparison to the complexities of many other calculations involved in submarine design. For instance, re-designing the duct intake of blade arrangments of a pumpjet to maximise efficiency at a different speed whilst simultaneously attempting to avoid cavitation would be vastly more complex and uncertain given the unstable difficult-to-model dynamics of turbulence and cavitation involved.  In contrast, scaling up a hull-diameter was quite simply and successfully accomplished when the Collins Class submarine was developed by scaling up both the length and diameter of it's smaller Swedish predecessor.  In 2006, DSTO commissioned a report which was dedicated to proposition of a different hull diameter being used for submarine used to replace Collins \parencite{joubert2006} in order to achieve lower hydrodynamic drag.  If such a development path was actually thought to impose  such significant challenges on the engineering and certification process as RADM Sammut's remarks imply, there is no way that in 2006 when the replacement was presumed to be undertaken for a smaller number of new submarines which would enter service in 2026 (as opposed to mid-2030s as currently planned) that such an undertaking would have been seriously considered.

That this ranks as a significant consideration in the mind of the chief of this program should be the cause of considerable alarm to those taking an interest the project, as it would appear that he is quite alone in this belief that this really is a significant or risky hurdle in a project of this scale.  Were it known to the competitors that their ability to achieve this simple engineering requirement was in question, one might assume they would be highly offended, since any contender that was not considered competent to design a pressure hull of a new diameter would presumeably never have been invited to participate.  RADM Sammut should be called upon to clarify what he meant by ``we needed to make sure that the certifying process would match the design philosophy" because if indeed he meant that Defence had already decided that their own certifying process already precluded scaled-up diameter hulls, this represents a very bold departure from what might be considered normal risk assessment processes for major engineering projects, including those used in designing and certifying the Collins.  It would also have essentially precluded both the Germans and Japanese from the outset.  Furthermore, if the advantage was considered so substantial, he should be called upon to substantiate how significant increased difficulty in certification would have been, with respect to cost and schedule, noting that the French proposal had a larger cost and longer budget than either of the other two contenders.

Amongst RADM Sammut's additional claims regarding the reason for choosing DCNS/Naval Group, include remarks regarding the establishment of 'sovereign capability' and 'data transfer' to Australia.  One would assume that this meant that the French had demonstrated a willingness to supply complete and timely information on key technologies, and that RADM Sammut's exisiting experience with the French supplying such information on the pumpjet (as Sean Costello had claimed had taken place during and prior to the conclusion of the CEP) supported this view.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{31 Oct 2017 - Greg Sammut, head of SEA 1000}]
\textit{In an interview that was published in a post\parencite{sammut2017} on the ASPI Strategist the leader of the future submarine project for the RAN responds to a question from Brendan Nicholson asking whether there is any doubt about using the pump-jet.}
\\
\\
``We are using pump-jet propulsion that was offered by France. Some claims about the pump jet are misleading. It's not purely for use in nuclear submarines. It can be efficient across the entire speed range, taking account of the submarine's size, the speed required and stealth. It needs to be tuned to the hull of the submarine. The hull needs to be tuned to the pump jet. We did not specify that the boat should have or not have air-independent propulsion or lithium-ion batteries, rather that it met our requirements."
\end{tcolorbox}

This comment confirms that the official policy is still to use the pumpjet, implying that the decision has indeed been made already.  This is at odds with the statement made only a few weeks earlier by Jean-Michel Billig.  This is also the first noteable reference to the inter-relationship between the hull and the pumpjet, implying that the pumpjet will indeed be modified in order to suit the hull, and that the hull will be modified to suit the pumpjet.  This contradicts the notion that the same pumpjet as France is fitting to the Suffren (i.e. nuclear barracuda) will simply be inlcuded 'as is' on the Australian non-nuclear 'Shortfin' version.  Given that the length and diameter of the hull were announced as being already decided and practically identical to the nuclear version, it raises the very serious question as to whether/why the changed shape of the hull would necessitate modification of the complex pumpjet.  As pointed out in the my testimony to the Senate \ref{SENmorrison}, the emphasis of this relationship, above the vastly different operational speed profile of the nuclear and non-nuclear submarines which would affect the propulsion system profoundly, raises serious questions about the origin of the comment.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{3 November 2017 - Department of Defence, in response to Question on Notice from Senator Xenophon from October 2017}]
\textit{As quoted in an ASPI Strategist blog post\parencite{davies2017}, and also by Senator Patrick in \ref{SENjohnson}, the Department of Defence argued in defence of the pump jet's efficiency.}
\\
\\
`A [pump jet] can provide higher propulsive efficiency across the speed range of a submarine, including the patrol speed of conventional submarines'.

\end{tcolorbox}

This corroborates that the comments made by RADM Sammut regarding efficiency were also made by the Department, under due consideration, to a Question on Notice.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=red!5!white,colframe=red!75!black, title= \textbf{15 Dec 2017 - Senate Estimates}]
\textit{Senator Rex Patrick asks Stephen Johnson (General Manager Submarines) about the pumpjet proposal, and produces documents suggesting they are less efficient at low speed.  The Defence Chief Scientist, Dr Zelinsky, and Chief of Defence Force, Air Chief Marshal Binskin, and Head of Capability and Sustainment Group, Kim Gillis are also involved in the discussion, defending the choice of the pumpjet.  Senator Patrick also specifically raises the concern about the impact of the weight of the pumpjet at the extremity of the submarine.   Full exchange in \ref{SENjohnson}.}
\\
\\
\textbf{Senator PATRICK:} Mr Johnson, there's been a lot of talk about pump jets. I've now tabled a document which has come from the United States, from Connecticut. You'd be very familiar with the organisation there. That basically describes the characteristics of pump jets, their pros and cons, as any piece of engineering equipment has. The document-and there's also another document that I've taken from an ASPI site-shows the efficiency characteristics of a pump jet. Both the US document-now declassified-and the document from the ASPI website suggest that, at low speed, pump jets are relatively inefficient compared to, say, a propeller. That may well be the reason why we see a pump jet on a Mark 48 torpedo but generally not on merchant ships and so forth. Would you care to comment on that?

\textbf{Mr Johnson:} As we said just a little earlier, the design process-whether it is a propeller or propulsion-is very much keyed to the criteria that you look for. The first thing I said was speed. \textbf{No matter what kind of propeller or propulsion pump jet you design, it has to be optimised for a particular speed.} There are classes of surface ships that look to optimise cross-ocean transit. Physics usually keeps you around 17 to 18 knots, and then there is the cost of maintenance and the cost of repair and the fact that there's no particular stealth requirement. Typically, you will find that those ships will use an open-tip propeller. \textbf{However, other ships, cruise ships in particular, use an electric shrouded pod, which is a variation of a pump-jet technology.} Of course, the technology gets changed over time, so the older a document is, the more it's dated to that point in time.
\\
\\
...
\\
\\
\textbf{Mr Johnson:} We were talking about optimising for the speed and the characteristics of the pump jet. The decision-making process for submarines is a little more complicated because there's an acoustic performance aspect as well as the pure efficiency of the blade-propulsor-pump jet arrangement. \textbf{Without going into details, it's a decision that we make with Navy, but I'm confident, having done this a few times already, that we are going to find this pump-jet approach to be the optimum for our particular circumstances for this particular submarine.}
\\
\\
...
\\
\\
\textbf{Senator PATRICK:} Okay. I'll go back to you, Mr Johnson. Noting that you're saying that this is based on something that hasn't been to sea yet - it will go to sea on the Barracuda; presumably that's the submarine you're referring to-there's still a risk that's sitting there. As the guardian of the taxpayer, I know where you sit. I have a responsibility from the taxpayers' perspective to ask: do you have a fallback position if the data that comes from the French trials, or other trials that you might do, presents you with data that's inconsistent with the operational profile that you intend for the submarine?

\textbf{Mr Johnson:} Yes. \textbf{We will understand the performance data early enough in the design that we can accommodate whatever is required. And then, just to reassure you, it's really going to be an adjustment to the design of the pump jet, not a decision to leave it.} Without going into too much detail, the curve that you're referring to is about the efficiency of the propeller, not about its acoustic performance. As you know, there's nothing free in physics, so those are different-shaped curves. What we seek to do, as Chief of the Defence Force said, is to make the right balance of multiple dimensions of this now.

\end{tcolorbox}

This testimony contradicts the assurance provided in an October interview with RADM Sammut.  It is clear from multiple parts of Mr Johnson's statement that the decision on whether to use the pumpjet or not has not yet been taken, since he is ``confident... that we are \textit{going to find} this pump-jet approach to be the optimum...".  Whilst acknowledging that he doesn't yet have the performance data from the nuclear Barracuda class (which is not yet in the water), he claims that it will be obtained early enough to ``accommodate whatever is required", which he gives assurance is only going to be adjustments to the design of the pumpjet.  This suggests that it is still not even known whether the pump-jet that was an `integral' part of the bid was offered in an unmodified state, or whether it was known by France that it would need to be modified from existing designs, or whether what was offered was actually only a skill-set and capability for the design of a new pumpjet.

In responding to the evidence produced by Senator Patrick that pumpjets may have an efficiency issue at low speeds, Mr Johnson refers to an `electric shrouded pod' which is used on cruise ships, which he claims is a modification of a pumpjet.  Nothing about this section of his testimony seems to have any reference of connection to reality.  Cross-ocean transits regularly occur at higher speeds than 17-18kt, including by cruise ships, which he mentions. These speeds are essentially unreacheable (or only extremely briefly sustainable) by a diesel electric submarine.  As pointed out in my evidence to Senate Economic References Committee \ref{SENmorrison} I can find no evidence of any cruise ships having any kind of shround around the tips of their propellers, or bearing any resemblance or relationship to a pumpjet.  In the same session, whilst in Mr Johnson's presence, it appears that Mr Johnson may have confused electric Azipods, which are used on large ocean-going ships such as cruise ships, and the Navy's new LHDs, with pumpjets.  Azipods have open-tipped propellers, and are not remotely related to pumpjets.  So far as I can discern, this particular part of the testimony was made-up entirely impromptu without any knowledge or care for the truth of the matter.  The degree of confidence that we can have in his assurance that the pump-jet will be the optimum approach for the submarine should be considered in light of this.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{9 February 2018 - Department of Defence, in response to Question on Notice from Senator Rex Patrick in December 2017}]
\textit{Senator Patrick inquires about the nature, cost, and conclusions of earlier research into pumpjets undertaken as part of \$214 million worth of ``de-risking" studies already undertaken, as well as the schedule and cost of subsequent research and development of the pumpjet.  Prior research consisted of a single 99 page report obtained via literature review and discussion with experts at a cost of \$68,825.  Ongoing research included \$1.1million spent to date to develop a generic submarine and pumpjet design, compuational and analytic models, and validation of these models through experimental studies.  The development and test program is not scheduled to commence until the last quarter of 2018.  Full questions and responses can be found in \ref{QON148}, \ref{QON150}, \ref{QON151}}
\\
\\
``The study concluded that pumpjets have the potential to deliver increased propulsive efficiency and lower shaft torque requirements, relative to an open propeller. This should lead to beneficial platform systems and operational impacts to increase vessel speed and/or submerged endurance.
The study developed a parametric model of pumpjet thrust and efficiency that has been used as the basis for more recent studies into pumpjet performance."
\\
\\
``Specifications for the pump jet are being developed as design of the Future Submarine proceeds. Initial inputs will be developed in the last quarter of 2018, which marks the beginning of the development and test program for the pump jet. Development, testing, and assurance will be conducted by several parties including Naval Group, Direction generale de l'armement (DGA), and DSTG. Testing and assurance activities will take place in France (including at the premises of the DGA Techniques Hydrodynamiques in Val de Reuil) and Australia. Testing will be largely completed by CDR."

\end{tcolorbox}

These responses represent another chance for the Department of Defence to articulate a considered opinion with regards to the merit of the pumpjet.  On this occasion, the mention of efficiency is accompanied by no specific claim with regards to the plausible speeds at which this might apply.  Instead, a mention of shaft torque (force per degree of rotation) requirements is made.  Given that pumpjets are known to be smaller in diameter and rotate at a higher speed, the requirement of lower torque is a trivial observation.

The statement that `inputs for the pump jet' will be developed in the last quarter of 2018 seem to imply that the pumpjet system is actually being designed ground-up, with its own `development and test program'.  Given that the hull diameter is known to be exactly the same \parencite{kerr2017}, and the length barely altered from the nuclear Barracuda, that the pumpjet is being altered at all on account of the hull, as directly implied by RADM Sammut \parencite{sammut2017}, is extremely surprising.  As discussed in my evidence to the Senate \ref{SENmorrison}, the prospect that a pumpjet even could be effectively `tuned' to match such a radically different operating profile with regards to speed is extremely odd.  I likened it to being ``as if someone bought an efficient hybrid car like a Prius and then decided to fit it with tank tracks from a dozer or a battle tank, and then argued they would tune the tank tracks back to be as efficient as the Prius was in the first place. It goes radically against what the technology was intended for in the first place."

Perhaps more remarkable is what this implies about the value (or eventuality) of pumpjet technology being transferred from France to Australia, and how that sits with statements made by RADM Sammut in particular.  DCNS originally boasted about their intention to transfer the technology to Australia \parencite{costello2016}.  RADM Sammut has since claimed that the way that ``data would be transferred to Australia" was amongst the reasons for his selection of DCNS as a design partner, whilst simultaneously citing the importance of the hull-diameter remaining identical to the nuclear submarine from which ours would be derived, in order to avoid ``a whole new set of calculations".  Only days later, he refers the need for the jet to be ``tuned to the hull" and the hull would be ``tuned to the pump jet'. In contrast to the `tuning' statement, this answer describes a ground-up development process from inputs which are yet to be determined, which raises a serious question as what value there is in having a design partner who intends to transfer that technology from near-complete submarine with an identical diameter hull.  One would assume that if such a ground-up development process were to be embarked upon which would necessarily have a feed-back loop with the final shaping of the hull, this would necessarily substantially heighten project risk. One would further assume that substantial work have been done to establish the suitability of the technology in advance, and the viability of reaching sufficient performance to meet the likely inputs in relatively tight development spiral with limited impacts on other systems or capabilities.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{3 April 2018 - Department of Defence, in response to Question on Notice from Senator Chris Ketter in December 2018}]
\textit{Senator Ketter asks Question 1, referencing specifically Greg Sammut in \parencite{sammut2017} and my paper \parencite{morrison2018} asks specifically whether Defence maintians that the pumpjet can be efficient across the entire speed range. Full question can be found in \ref{QON1}.}
\\
\\
Question: Does defence still maintain that a pump-jet can be efficient across the entire speed range?
\\
\\
Answer: ``The efficiency of pump-jets designed for submarine use cannot be determined through comparisons with water-jets and ducted propellers, which are different in their operation from pump-jets. Moreover, the global efficiency of submarine pump-jets cannot be derived from `open water' tests of pump-jets, where the pump-jet is assessed independently of the hull of the submarine. A correctly designed pump-jet, suited to the aft hull shape of the submarine, achieves better optimisation of efficiency and stealth than any propeller, even at low speeds."

\end{tcolorbox}

This answer to a question again rests on the claim that the influence of the hull means that comparisons and studies such as those referred to in \parencite{morrison2018} can be ignored.  This is a question which will be given a substantial reply in subsequent sections, but even a quick reading superficial reading reveals alarming levels of over-reach in the answer.  (For intance, a type of ducted propeller known as a decelerating ducted propeller actually \textit{is} a pumpjet, by definition. An entire chapter (5) of the paper to which Defence is supposedly responding \parencite{morrison2018} discusses this at some length.)

However, the response also avoids directly answering the question which specifically refers to efficiency in isolation.  By only claiming that the pumpjet ``achieves better optimisation of efficiency \textit{and stealth}, even at low speeds", the response avoids directly contradicting the evidence that pumpjet efficiency inevitably declines dramatically at low speeds, which is the core contention of the paper in question.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{3 April 2018 - Department of Defence, in response to Question on Notice from Senator Chris Ketter in December 2018}]
\textit{Senator Ketter asks Question 2, inquiring as to how Defence assesses Naval Group's opinion regarding the pumpjet's suitability, following Jean-Michel Billig's comments.  Full questions and responses  can be found in \ref{QON2}.}
\\
\\
Question: Is Defence still confident that Naval Group still hold the opinion that the pump-jet is well suited to diesel-electric submarines?
\\
\\
Answer: ``Defence intends to use the pump jet propulsion system for the Future Submarine that was offered by Naval Group as part of their Competitive Evaluation Proposal and remains on offer to Australia."

\end{tcolorbox}

This constitutes a point-blank refusal to answer the question that was asked.  Given the earlier carefully worded comments of the most senior Naval Group officials, and careful placement of ultimate responsibility for the choice with the Commonwealth, such a refusal to answer a question is alarming.  If the opinion that the pump-jet was overall well suited to diesel-electric submarines was held by Naval Group, it would be only natural to expect that that would be confidently communicated to the Department of Defence, given that they had assessed their bid which it was `integral' to.  Provided that opinion was communicated and hadn't changed, it would be easy and expected that Defence would give a confident answer to the question. If Naval Group offered a proposal which included technologies were known to be poorly suited together would constitute at the minimum a severe breach of trust, and quite probably fraud.

Overall, this response casts serious doubt on the type of communication and advice which has flowed between Naval Group and the Department of Defence regarding the pumpjet to date.  As a ship-builder which had designed and manufactured both pumpjets and propellers, and fitted them to both diesel-electric as well as nuclear submarines, there could be no excuse for Naval Group to not have a well-informed opinion on this matter.  For Defence not to have obtained that opinion from DCNS at the time of considering the bid would be a matter of incomprehensible oversight.  It is made all the more difficult to believe given the scrutiny applied to the matter in the Senate in 2012 \ref{SENmoffit}, and the program director's assurance that the work was being done to assess its suitability.

Naval Group certainly said that the performance of the pumpjet was known to the Government prior to the close of the CEP \parencite{costello2016}.  It is a matter of substantial public interest to determine whether that performance information given at the time was sufficient to communicate Naval Group's opinion on the suitability of the pumpjet to a diesel-electric submarine, or allow Defence to confidently form their own opinion based on this information. If Defence came away with a positive opinion of the pumpjet's suitability whilst Naval Group actually privately maintained a negative one, (and elected not to avail Defence of their false impression for commercial reasons) this would amount to a gross betrayal of trust and quite possibly fraud on the part of the Naval Group and the French Government. It would also demonstrate an catastrophic level of incompetence on the part of Defence, given their stated efforts to assess the pumpjet prior to this point.  If Defence has since become aware of Naval Group's lack of confidence in the pump jet's suitability, this may explain the defence response to this question, which explicitly refers to the pump jet propulsion being offered `as part of their Competitive Evaluation Proposal' and `remains on offer to Australia', which could be read to deflect responsibility for technology choice to Naval Group, implying they would or ought to withdraw the offer if it was now known to be inappropriate.

In such a scenario Defence is also acting to shelter Naval Group from accusations of fraud as well as the likely legal proceedings that would be open to the Commonwealth if information provided in the CEP was found to be fraudulent.  It's probable that this constitutes misleading parliament, but of far greater concern is the possibility that the necessary corrective action to avoid a procurement disaster of unprecedent scale is not yet being taken by Defence.  Why that could be the case is also a matter of substantial public interest.

It is possible to believe that Defence's own embarrassment at having been unable to detect the insincerity in Naval Group's offer of the pumpjet, and inability to reach their own technical conclusions quickly enough to detect the fraud, may motivate the present actions.  This amount to a disasterous failure on the part of the public service to act with integrity and in the national interest.  However, if Defence had truely been un-wittingly deceived, it is difficult to see how they could not recover their pride (and regain a necessary measure of respect from their build partner) by quickly adopting a more combative stance, calling out the deceptions on the part contractors part, and quickly exacting whatever contractual reparations they might be entitled to.  If the contractual reparations were not sufficiently strong, the Commonwealth's freedom to not proceed to a build contract (from the initial design contract) or re-introduce competition into the design process as many have suggested should occur in any case \parencite{stanford2017}, would have provided ample commercial leverage to reset the relationship in a manner which was better aligned with the Commonwealth's interest.

One must at this point give consideration to the possibly that Defence had not been entirely unwittingly deceived, and was actually complicit in their own deception, and it's present upkeep. This would mean that the Department of Defence and the Navy, or at least a few of very senior people within those organisations, had to conspired allow a proposal which was known to be inappropriate to the Commonwealth's needs to appear victorious in the CEP,  presumeably in order to favour a bidder for whom no more defensible explanation could be found.  It appears unlikely that there could be any decent motive for such a course of action.  However, one must evaluate the probabilities of the necessary oversights for such an eventuatlity occurring by accident or by arrangement.  If the pumpjet was known by DCNS to be an unsuitable proposition for a diesel-electric submarine, a seemingly unlikely degree of naive incompetence would be required for Defence to fail to ascertain as much during the Competitive Evaluation Process.  To remain genuinely unwitting for two subsequent years of design development would seem even more so.  The possibility of some corruption influencing the selection of Australia's \$50bn future submarine contract must at least be considered.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{3 April 2018 - Department of Defence, in response to Question on Notice from Senator Chris Ketter in December 2018}]
\textit{Senator Ketter asks Question 3, which focuses specifically on whether Defence believes the pumpjet is acoustically superior across the entire speed range, or just at higher speeds, referencing my paper \parencite{morrison2018}, Herve Guillou \parencite{guillou2017} and Costello and Autret \parencite{autret2016}.  Full question and response can be found in \ref{QON3}.}
\\
\\
Question: Is Defence of the belief that the pump-jet is quieter than propellers at all operating speeds, or is it the case, as suggested by Mr Morrison, that at some low speeds propellers are acoustically superior, and only at higher speeds is the pump-jet acoustically superior?
\\
\\
Answer: ``See response to Senate Economics Reference Committee Question on Notice, Q1 – Pump Jet Efficiency Across the Entire Speed Range."
\end{tcolorbox}
The response to Q1 - a question specifically and only asking a question about \textit{efficiency} of the pumpjet, avoids contradicting the evidence directly by conflating the question with acoustics.  To avoid the answering a question that is specifically about acoustics by referring to that same answer which incorporates efficiency is nothing more than a point-blank refusal to answer either question directly.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{17 May 2018 - Brent Clark, Acting CEO Naval Group Australia}]
\textit{In an interview that was published in a post\parencite{guillou2017} on the ASPI Strategist the Autralian CEO of Naval group explains to Brendan Nicholson why Naval Group is recommending the pumpjet for the Shortfin Barracuda, and answers some critics.}
\\
\\
The company building the Navy's 12 new submarines says they'll be the world's first conventional boats big enough to effectively use the company's pump jet propulsion system. French Naval Group Australia says that its propulsor will prove much quieter than a propeller on the 4,700‑tonne submarine it has dubbed the 'Shortfin Barracuda'.
\\
\\
Naval Group Australia's interim CEO, Brent Clark, tells The Strategist that the decision about the best propulsion system for the submarine will be made by the Defence Department, but that he hadn't had any indication that it would chose anything but a propulsor. 'We've had every indication that they like what was offered.'
\\
\\
The propulsor that Naval Group is recommending for its Shortfin Barracuda consists of a propeller within a duct to reduce noise and the water disturbance known as cavitation, which can give away a boat's position to hunters. Clark says he's read many negative comments about propulsors, but some critics used very old studies to try and justify their argument.
\\
\\
"A lot of people don't fully comprehend, I think, that yes, a propulsor on a 1,500-tonne coastal submarine would make no sense. But we're not building a 1,500-tonne coastal submarine. We're building quite a large submarine. And this is the first time that a conventional submarine's been large enough to take a propulsor."

\end{tcolorbox}

Five months after the previous Senate hearing, this blog post confirms that the final choice on whether the pump-jet is to be included is still yet to be made, and Naval Group is still clear that the decision rests with Defence, rather than being assumed as tied to their bid.

This is also the first time that the argument that the size of the submarine is a crucial enabling difference that makes the pumpjet viable is made.  It is the closest that Naval Group has come from making a firm statement of faith in the suitability of the technology for the Australian diesel-elctric submarine. As I discussed in another blog post in ASPI Strategist \parencite{morrison2018b} shortly afterwards, this is a remarkeably concession that is consistent with the physics that I outlined, since the inefficiency of the pumpjet would likely drive demand for additional fuel and batteries (hence size) to meet requirements, which is more of a concession of the detrimental impact of the pumpjet on the submarine.  This particular blog post was acknowledged on twitter by one former US Navy submariner and submarine enthusiast as an ``Awesome article about pump jet efficiencies and viability".  Naval Group was tagged in the re-tweet, which was subsequently re-tweeted and liked by naval experts, publications, commentators and authors from around the world.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=red!5!white,colframe=red!75!black, title= \textbf{07 Jun 2018 - Senate Economic References Committee}]
\textit{The author presents evidence to the final public hearing of this committee's Inquiry into the Future of Australia's Naval Shipbuilding Industry, challenging the pumpjet choice.  Brent Clark, the outgoing CEO Naval Group Australia, Stephen Johnson, the General Manager Submarines from Defence, as well as Kim Gillis, Head of Capability and Sustainment Group at Defence, are all invited to respond to author's evidence.  Full extracts of all discussion on the pumpjet topic are in  \ref{SENmorrison}.}
\\
\\
\textbf{Mr Morrison:} ...In my report, I made the conclusive comparison that, between two otherwise identical submarines, the one with the pump jet will always have a lower dived endurance, a lower dived range, a worse indiscretion ratio, a lower overall endurance and a lower overall range than the one with the propeller. This statement, so far as I know, has yet to be challenged, and \textbf{I'd be very surprised if anyone with technical competence or professional integrity sought to challenge it.}
...

I'd like to look at one further public statement that was made. That is from Stephen Johnson. This was made in relation to an answer that was given to this committee some time ago. When challenged very specifically about efficiency by Senator Patrick, he made this statement:

``No matter what kind of propeller or propulsion pump jet you design, it has to be optimised for a particular speed. There are classes of surface ships that look to optimise cross ocean transit. Physics usually keeps you around 17 to 18 knots, and then there is the cost of maintenance and the cost of repair and the fact that there's no particular stealth requirement. Typically, you will find that those ships will use an open-tip propeller. However, other ships, cruise ships in particular, use an electric shrouded pod, which is a variation of a pump-jet technology."
...
\textbf{I wonder whether here Mr Johnson has actually been confused with the Azipods that are installed on large cruise ships. They do use an electric motor, that's the only similarity I can find with the pump jet, and are indeed open to propellers and are quite similar to the propulsion systems used on the LHD. That is the best explanation that I can find so far for these comments. I certainly can't find any evidence of any shrouds being used on cruise-ship propulsion. Moreover, I can't find any evidence of pump jets of this sort being used on any ship, for which economic travel is a main priority, operating below the speeds of about 25 or 30 knots. So it's very interesting to wonder what exactly Mr Johnson was referring to in this case.}
\\
\\
...
\\
\\
\textbf{Senator PATRICK:} Firstly, it's probably proper that we give you an opportunity to respond to Mr Morrison's evidence. That's probably proper, Chair.

\textbf{Mr Clark:} Thank you, Senator Patrick. I believe this is the first time we've sat opposite each other in a committee, so let the fun begin! From my perspective, \textbf{I don't intend to offer any commentary on the previous speaker.} Suffice to say that \textbf{the information that we've been supplying of a classified nature to the Australian government is held by the Australian government and was supplied by the French government.} The Australian government and the Australian Department of Defence obviously have access to classified research from the United States Navy, the Royal Navy and the French Navy on this matter, and I don't intend to get into a debate on this topic.
\\
\\
\end{tcolorbox}

Naval Group here steadfastly refuses to offer a clear opinion on the pumpjet. Only months earlier Defence was asked a specific question on notice about whether they thought Naval Group maintained that a pumpjet was suitable for a diesel-electric submarine.  Clarke's interview on the Strategist, which seemed to be strongest statement of faith from Naval group in the suitability of the pumpjet for a conventional submarine (perhaps in an effort to patch up the noteable refusal to answer the question) had been answered publically by the same person who had just appeared and he was being asked to respond to.  It might have been expected that Mr Clark would at least attempt to demonstrate conviction about the merits of the proposal.  Instead, he described the information that they were supplying as a classified, government-to-government exchange, as if he were an impartial bystander.  No assurance of the proposals superiority, suitability, or viability in any respect was made.

Observers in the room or viewers of the video record may also observe that Mr Clark seemed to cut his statement short quite spontaneously at the point when he reached descriptions of engineering accomplishments surrounding the propulsion system, and there is no apparent record of the remainder of his statement actually being tabled as he said he would.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=red!5!white,colframe=red!75!black, title= \textbf{07 Jun 2018 - Senate Economic References Committee - CONTINUED}]

\textbf{Senator PATRICK:} Sure. I have one more question, and this is to give you an opportunity to respond to Mr Morrison's evidence, Mr Johnson. He contradicted your evidence about cruise ships having pump jets. \textbf{He said that you might have meant Azipods?}

\textbf{Mr Johnson:} \textbf{I don't remember that clearly enough.} I do remember that you had produced a document, and I was trying to answer questions without the benefit of where we were going.

\textbf{Senator PATRICK:} Just giving you an opportunity to-

\textbf{Mr Johnson:} I don't have a clear idea of exactly the choice of words I used, but I think the concept remains the same: propellers and propulsers are designed for very, very specific purposes, and if you change the purpose you will change your choice of design.

\textbf{Senator PATRICK:} Thank you.

\textbf{Senator KIM CARR:} Mr Johnson, you sat through the evidence this morning. I noted you at the back of the room when Mr Morrison made his presentation-is that correct?

\textbf{Mr Johnson:} \textbf{Yes, I was here this morning. I wasn't necessarily paying attention.}

\textbf{Senator KIM CARR:} You weren't? That's a big thing to admit on the Hansard!

Senator Reynolds interjecting-

Senator Kim Carr interjecting-

\textbf{Mr Gillis:} I will support my colleague: not all of this is riveting!

\textbf{Senator KIM CARR:} Is that right?

\textbf{Senator REYNOLDS:} Mr Gillis!
\\
\\
...
\\
\\
\textbf{Senator KIM CARR:} I contend that we don't have the technical expertise to evaluate the report. It may well be that we recommend that further work be done on it, but I'd want to see your response to this-to give us a critique of it; your perspective as to the report. He's made some assertions to the committee. You no doubt will get the Hansard pretty quickly, if the secretary can make sure that that's available to the officers.

\textbf{Mr Gillis:} The only caveat I'll put on any response we make is that the one thing we do really well is detailed engineering analysis. When you have somebody's report-\textbf{and I have not read the report yet; we weren't actually given a copy.} To actually go through it-and it looks like something which is yea thick-with serious analysis is going to take us time. We can only do what we can do.

\textbf{Senator KIM CARR:} If anyone comes to me and says that fundamentally you've got it wrong on a technical matter such as a propulsion system, (1) we have to take notice of that, but (2) I'd be surprised, frankly, given the level of technical expertise available.

\textbf{Mr Gillis:} Senator, that's one. I regularly get correspondence from the public saying that almost every decision that we've made for the last 50 years on a defence matter is wrong, because somebody has an opinion.

\textbf{Senator KIM CARR:} Not since the fall of Singapore, surely!

\textbf{Mr Gillis:} Almost since the fall of Singapore. \textbf{There are many people out there with opinions, and the vast majority-I would say over 99 per cent of them-you can dismiss within the first page of reading them because they are fundamentally wrong.}

\textbf{Senator KIM CARR:} We clearly don't do that, because Mr Morrison wouldn't have got a hearing at this committee if we'd dismissed it. I'd like an assessment from you.

\textbf{Mr Gillis:} I accept that. We will provide the committee with the best possible advice we can in the shortest possible time frame to meet your requirements so that you can be better informed, because we do hold that expertise.
\\
\\
...
\\
\\
\textbf{Mr Johnson:} Without contemporary knowledge or without understanding necessarily the intricacy. For example, the shaft horsepower-the power required to move even a very big ship at very low speeds-is astonishingly small. If you look at inefficiencies, you look at a curve and say, 'That must be a problem.' It's not. It's like: 'Hey, you paid twice what you needed to for that. \textbf{It was a penny. It cost you two pennies. That's 100 per cent over.}' So some of this is out of context with what we're doing, which will make it hard to give you an answer of empirical-

\end{tcolorbox}

This exchange demonstrates a dismissive attitude towards the evidence I had presented, which borders on contempt. The attitude is clearly pervasive and supported in the higest levels of the defence department.  The General Manager Submarines, Stephen Johnson, claims he simply cannot recall what he said about pumpjets being related to pods on cruise ships, despite being physically present in the same room when I pointedly quoted and contradicted him.  This being pointed out, he claimed to simply not to be paying attention, which won the support of the Head of Capapbility and Sustainment Group, Kim Gillis.  Despite the obvious shock and exclamation of the senators, Mr Gillis went on to attempt to dismiss my evidence as just another opinion which could probably to be dismissed along with 99\% of the others.

This level of ambivalence towards external criticism from a such a senior official on such a topic as this should be alarming.  The topic of the pumpjet had such sustained criticism and interest from various quarters, including multiple questions from senators from all sides of government over six years, reports in the ABC, ASPI Strategist, Defence Connect, Insight Economics, and many others.  My paper was commissioned by a concerned citizen who wanted to have a better investigation into the physics underlying the controversy which had already received considerable attention, albeit with confusing results.

Such an attitude raises questions about whether Mr Gillis's evidence is even given honestly.  For instance, Mr Gillis' claim that he had not read the report, since ``we weren't actually  given a copy" warrants serious attention.  My paper was made publically available with the blog-post summarising it on the 14th of February, and as well as in my reply to Brend Clark on  the 24th of May.  Moreover, his department had received written Questions on Notice from Senator Chris Ketter, acting on behalf of this very committee, specifically requesting written responses to my paper with a return date of 3 April.  His department did indeed responded to prior to the hearing, as noted above with responses which blatantly refused to answer the questions. Furthermore this particular hearing was also especially long in the making, since it was once postponed on account of Defence refusing to make officials available to attend.  My presence as a witness was requested in February when the proposed date was still in May, and the list of witnesses was provided publically circulated during the weeks prior to the hearing.

In such circumstances, claims to be ignorant of my report, and an attempt to lump it with the other 99\% of dissenting correspondence, seem about as reasonable as Mr Johnson's claim that he simply wasn't paying attention when he was being quoted and refuted before the committee earlier in the day.  That this is the type of responses being given by the most senior officials who are responsible for the largest defence procurement in Australia's history should be alarming.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{August 2018 - Dr David Kershaw, in Australian Defence Magazine}]
\textit{Julian Kerr includes extensive quotation from Dr David Kershaw in an article titled "Future Submarine: Pumpjet propulsor technology choice still not well explained" \parencite{kerr2018}.  Dr Kershaw is the current Chief of Maritime Division, a position he held since May 2016.  He was previously the Science and Technology Advisor for SEA 1000 between July 2011 and July 2014.}
\\
\\
As explained by Dr Kershaw, there was a degree of misunderstanding concerning pumpjet design and performance.  To propel a surface ship or submarine involved overcoming resistance to push the vessel through the water with a suitably-designed propulsor.
Surface ships need to overcome wave-making resistance, which varies with speed.  Their propulsors therefore need to be optimised for the required operating speed; when operating at off-design speeds, efficiency drops off.
\textbf{Since submerged submarines do not have to overcome wave resistance, the efficiency of a propulsor can be considered independent of speed} and the aim is to design an efficient propulsor.
In other words, selecting the propulsor for a submarine is driven by the types of mission it is expected to undertake and where the boat will operate, Dr Kershaw stated.
\\
\\
``The question of propeller versus pump jet really comes down to the acoustic signature; a well-designed pump jet is considerably quieter than an equivalent propeller." he said. "One of the advantages of a well-designed pump jet is that it can smooth out the water flow that is coming off the back of a submarine before the flow hits the moving rotor and there's less unsteady flow, and thus less noise.
``This smoothing-out also increases the propulsive efficiency; a well-designed pump-jet can have efficiencies equal to, or better than, an equivalent propeller.  You just have to have the capabilities and facilities to be able to design an efficient pumpjet and we know that France can do that."
Echoing RADM Sammut, Dr Kershaw stressed that the pumpjet on a submarine must be designed in conjunction with the hullform to which it was to be attached.
``Its' not like a propeller where you can design your propeller open loop i.e. without considering the interaction with the hull, and then attache the propeller to the hull.  \textbf{The water flow over the hull to the pump jet is so integral you've got to design it as a pair, and the hull and the pump jet have to be tuned together}, it's a total system that you're building."
\\
\\
``One of the issues with pumpjets and the reason they're not used on diesel-electric submarines is they're much heavier than a conventional propeller, although they have a smaller diameter.
\textbf{``As the size of the submarine increases} (the Future Submarine's 4,700 tonne submerged displacement is significantly heavier than the 3,407 tonnes of the Collins-class it will replace) \textbf{you start to get to the point where a pump jet can become sensible."}
\end{tcolorbox}

The first feature of this passage is Dr Kershaws's telling remarks about wave-making resistance, and how in the special submarine scenario the efficiency of a propulsor can be considered independent of speed.  This represents a pure, brazen, and obvious falsehood, justified by a seemingly-relevant truth.  It's akin to arguing 'camels have to cary two heavy humps on their backs.  Horses don't have to carry these two heavy humps, which is why they can fly.'  It also directly contradicts the claim made by Stephen Johnson before the Senate, that ``No matter what kind of propeller or propulsion pump jet  you design, it has to be optimised for a particular speed.".

Whilst it is true that deeply submerged bodies do not need to overcome wave-making resistance, all bodies in contact with any fluid must also overcome skin friction, wich means that resistance from this source scales with \textit{velocity squared} in almost all circumstances, including for ships.  This is a most fundamental and basic assumption which is absolutely intrinsic to all basic marine propulsion, and should also be understood by essentially anyone who has passed first-year engineering or physics courses.  (It's known) (For instance, see Pages 9 and 10 of Reference \parencite{MAN2017}, or Chapter 12 of Reference \parencite{Carlton2007}, or Chapter 3 of Reference \parencite{Mollard2011}.)

Apart from it being quite an elementary law of phyiscs (the Drag Law), referred to and relied upon in posts such those by Andrew Davies \parencite{davies2017} commenting on this exact question, any substantial reading of the literature focussed on the design of pumpjets for fully submerged objects (such as torpedoes or submarines) will also make it abundantly clear, and the special distinction with wave-making resistance (which is crucial for fast-vessels in the process of reaching the plane) is also abundantly clear, and has been well known for a long time \parencite{wislicenus1973} \parencite{henderson1964}.

That the Science and Technology Advisor on SEA 1000, and current Head of Maritime Division at DST could make such an gross error is extremely shocking.  The lack of understanding demonstrated strongly contradict the assumption that he is competent to speak or advise on this topic.  Whilst his expertise is most probably actually not that closely related to propulsion systems or hydrodynamics (his online bio suggests his torpedo-related experience is actually closer to the computer systems and signal processing involved \parencite{DSTG2018}) it would seem extremely strange for a competent physicist to such public statements outside their area of expertise, if they weren't confident in their own understanding.

The other revealing thing about Dr David Kershaw's remarks here is the clear connection with comments made by RADM Sammut.  He gives the most detailed description of the interaction between the hull and the pumpjet thus far, arriving at the same conclusion that they have to be 'tuned' together which has been previously used by RADM Sammut.  (This argument will be addressed fully in a later chapter.)  An apparent conclusion is that RADM may have relied upon advice from Dr David Kershaw, (which given the ineptitute of earlier comments relating to hydrodynamics is gravely concerning), or that both of them have relied upon advice from another common source.

The same conclusion might be reached here regarding the comments about submarine size, and a connection to comments made by Brent Clark for the ASPI Strategist \parencite{clark2018}.  The comments by Dr Kershaw cast the argument in a very different light, suggesting that it is the weight of the pumpjet itself (as opposed to the weight of the extra fuel and batteries it demanded) was what made the pumpjet only viable for a very large diesel-electric submarine than anyone previous had built.  This argument will be dealt with in a subsequent chapter.  However, if it is indeed the case that the former CEO of Naval Group Australia, Brent Clark, was relying on Dr Kershaw for advice regarding reasons why the pumpjet was viable for the Australian future submarine, this raises the question as to why Mr Clark doesn't have his own advice from within Naval Group.  This further throws into doubt that Naval Group still do hold the opinion that the pumpjet is well suited to a diesel-electric submarine.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{August 2018 - Dr David Kershaw, in Australian Defence Magazine - CONTINUED}]
Although pump jet propulsion would be new to the RAN, the technology was well-known to DSTG.
"We've invested heavily in our hydrodynamics and hydroacoustics program over the past 20 years and this is includes the computational models as well as experimental facilities that confirm modelling results," Dr Kershaw said to \textit{ADM}.
\textbf{We've been able to develop our analysis on a generic pumpjet design and thus confirm our understanding of pumpjet performance; you basically build it up from first principles.
"Our analysis at the moment is all around analysing current options but we will be looking at what's ahead in the future and enable Australia to develop the sovereign capability to take best advantage of pumpjet technology."}
\\
\\
So will pump jet technology definitely be used for the Future Submarine?
``That is what is currently in the design \textbf{and that's what a lot of our science and technology program is about; having the capability to do the independent analysis on the technologyies} and capabilities planned for the Future Submarine, including the pump jet program," Dr Kershaw said to \textit{ADM}.
\textbf{Discussions were currently underway with Naval Group and the French government regarding closer collaboration on a range of technologies and this would include pump jets, Dr Kershaw disclosed.}
The program was still being worked on, and on the Australian side would involve a government-only team to comply with French security regulations.
``One thing I would like to emphasis is that Australia's geographic circumstances require a larger conventional submarine than those currently produced for export to other parts of the world.  The size of the Future Submarine means a pumpjet is a real game changer for optimising the performance of the submarine in its intended role, and ensuring regional superiority in undersea warfare."
Dr Kershaw concluded.

\end{tcolorbox}

These comments are extremely important because the illuminate the true state of DSTG's research, and the degree of confidence they have in the pumpjet.  It is consistent with responses given to a question on notice from Rex Patrick \ref{QON150}, where it was revealed that 1.1 million was spent on ``development of a \textit{generic submarine and pumpjet design}, computational and analytics models, and validation of these models through experimental studies".  The admission that DSTG is still working on a 'generic' design also casts into new light their repeated insistence that the performance of a pumpjet cannot be assessed unless it is exactly tuned carefully with the hull, and vice versa.  It is plainly obvious that if they are still working on 'generic' models, rather than the particular one which is proposed by France, they are in no better position to make positive assessment that contradicts findings in Reference \parencite{morrison2018}.  By their own assessment, according to the insistence of Dr Kershaw and RADM Sammut, they are in no position to assess the performance of the pumpjet that will be used on the future submarine, given they haven't been able to 'tune' it to the hull shape of the future submarine, and 'tune' the hullshape to the pumjet.  They are still working with 'generic' designs.

As I have already put before the Senate \ref{SENmorrison}, I find that idea that such 'tuning' could so dramatically alter the performance of the pumjet to be completely implausible.  I would support the idea that the kind of generic modelling described would be quite sufficient to reveal results that are consistent with the results I produced from a basic literature review (that pumpjet efficiency declines towards zero at very low water-speeds) and safely rule-out the suitability of a pumpjet for diesel electric submarines. If the results of the modelling done by DSTG to date could confidently contradict such a finding, one would expect that in Senate hearings and in responses to Questions on Notice, a very confident and clear-cut refution of the claims made in my report and evidence would be able to be made.  This has been conspicuously absent from all responses, including this one.  Instead, it is claimed that the 'tuning' with the hull, nullifies the result. If instead, the DSTG studies are actually consistent with my findings, then DSTG and Defence are knowingly supressing such results, or naively hoping that the French design, and the finer 'tuning' between the pumpjet and the hull can avhieve such a dramatically diferent result that the results of 'generic' tests can be ignored.  Either situation would have terrible consequences for the program, and if the former is true it is likely that the Parliament has been deliberately misled on multiple occasions.

These comments also illuminate a complete failure on the part of Naval Group to supply sufficient information to the Australian government regarding the pumpjet. Instead of relying on Naval Group's own modelling of the pumpjet that is being fitted to the nuclear Barracuda (which will have the same hull diameter as the Australian Shortfin Barracuda according to RADM Sammut), or even the proven, in-service pumpjet used on the larger French SSBNs, the Australian Defence Science and Technology Group is having to develop their own analysis on a ``generic pumpjet design" to confirm their own understanding through a process which they ``build up from first principles".  The non-existent state of information transfer is further confirmed by the admission that \textit{discussions} are ``currently underway" for closer collaboration on this technology.

That DSTG is still working up from first principles to develop the capability to do ``independent analysis on the technologies" that we have accepted in a \$50 billion dollar bid from a supposedly trusted design partner should come to all observers as a shocking revelation.  It is a deeply humiliating position for Australia, as a customer and an ally, to be so denied access to sufficient information to clearly and confidently understand the suitability of this technology.

It is particularly shocking in light of the fact that the program head, RADM Sammut explicitly told Senate Estimates that, alongside the pumpjet, ``the way in which data would be transferred to Australia, the manner in which intellectual property would also be managed in the program" was part of the reason the partner was selected.

It is even more shocking in the light of the fact that the CEO of Naval Group Australia, Sean Costello's, publically boasted about the significance of the ``offer from France to transfer to Australia sovereign control and use of pump-jet propulsion technology" \parencite{costello2016}, and his simultaneous claim (prior to the conclusion of the CEP) in the same that the performance data of the pumpjet are known by Australian government.

It is still more shocking in the light of the then outgoing acting CEO, Brent Clark's further confimration before a Senate Inquiry \ref{SENmorrison} ``we've been supplying" classified documents on this topic to the Australian Government.

In contrast to the assurance of full technology transfer, the comments by Dr Kershaw in this article reveals the true state of information flow from DCNS.  Apparently, there is still not sufficent information to be confident about the use of the pumpjet.  Still, a confident decision on whether the pumpjet will even be used cannot be made. Consistent with the statements of Mr Johnson \ref{SENjohnson}, the information about the pumpjet performance of the pumpjet on the nuclear Barracuda has not yet arrived.  Nor, it would appear, is sufficient theoretical modelling, or scale model testing, or any other confirmatory data sufficient on this pumpjet for DSTG to be clearly persuaded of the suitability of the pumpjet for our submarine.  Whilst simultaneously claiming that efficiency and performance of the pumpjet can't be assessed without the precise form of the hull and the pumpjet being 'tuned' together, they admit that they are having to model a 'generic' pumpjet, building up from first principals, in order to have sufficient confidence in their own 'indepenent analysis' of the technologies.

It can be concluded firmly from this interview that the promise of technology transfer promised by Sean Costello has not been made good, even to a sufficient extent to allow the government to confidently decide on its suitability for the submarine, and assess Naval Groups confidence in the suitability sufficient to answer parliamentary Questions on Notice.  Such gravity of such a failure is of profound consequence to the well-being of such a major project, given the significance of the pumpjet to the overall weight, balance and performance of the submarine as I outlined before the Senate \ref{SEMmorrison}.

It also throws into doubt the integrity of the statements made by RADM Sammut before Senate Estimates.  In the days and weeks immediately prior to his testimony \ref{SEMsammut}, Jean-Michel Billig had triggerd the controversy around the pumpjet at PACIFIC 2017, the largest Naval Show and conference in Australia, which occurs only each two years. This had lead to substantial coverage on the ABC, Defence Connect, and ASPI Strategist, and almost certainly contributed to the line of questioning that Senator Hanson doggedly pursued about the pumpjet that day.  An interview where he faced further questions on the pumpjet's inclusion with Brendan Nicholson was published just days later\parencite{sammut2018}.  It is inconceivable that the significance of the pumpjet question, and the state of the Commonwealth's confidence in its inclusion in the design was not front of mind at the time when he was being questioned by Senator Hanson on the 25th of October 2017.  And he still chose to include the ``the way in which data would be transferred to Australia" as a reason for selecting the French, alongside the pumpjet, to deflect Senator Hanson's suggestion that the pumpjet alone was the main reason for selecting the French.

If Naval Group's level of information transfer was proceeding as promised, then RADM Sammut had accepted (as being superior to other bids proposals) a sequence of data and technology transfer which would leave DSTG relying on their own 'generic' models in trying to assess the suitability of the pumpjet, and Defence unable to answer confidently whether Naval Group even had confidence in the suitability of the technology two years after the conclusion of the evaluation process.  Unless every other competitor had also promised to withold critical information to assess the suitability of a technology that was 'integral' to their bid, there can be no plausible defence of such a decision as competent.  Alternatively, if Naval group had promised more fluid information transfer, and this information had subsequently been withheld, RADM Sammut was actually concealing that a crucial break-down in relations with the major contractor had already occurred, with potentially disasterous consequences for the future of the submarine program. The only other viable alternative is that information transfer was not really something that commonwealth looked at in considering the selection of DCNS as a design partner, and RADM was knowingly misleading the Parliament.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{August 2018 - RADM Sammut, in Australia Pacific Defence Reporter}]
\textit{Kym Bergmann includes extensive quotation from RADM Greg Sammut in an article titled "SEA 1000 Meeting Major Milestones" \parencite{bergmann2018}.  The following is an extract on the technologies relevant to propulsion.}
\\
\\
Sammut confirmed that \textbf{the current thinking is
for a pump jet} system to drive the submarines
through the water. Navy is of the view that
there are different characteristics to propulsion
systems using propellers, ducted propellers
and pump jets. \textbf{The performance of a pump
jet needs to be assessed in conjunction with
the characteristics of the submarine hull} and
not in isolation, because they are matched -
something that is overlooked in much of the
current literature. \textbf{Pump jets are also stealthier
than propellers across the entire speed range of
the submarine} - something particularly important
for the RAN because the submarines have to
remain undetected during very lengthy transits
and not only during the time on station.
\\
\\
An advantage that Naval Group brings to this is
that they have a great deal of experience with pump
jet technology, including for the highly successful
French SSBN fleet. This activity is also being
supported by our Defence Science \& Technology
Group (DSTG) and will result in scale models
being tank tested in France at a highly classified
facility that the author visited in 2017. On the other
hand, \textbf{the first batch of submarines will use lead acid
batteries rather than lithium ion because the
RAN does not want to be introducing a massive
number of new technologies simultaneously.}
\\
\\
The Navy \textbf{is continuing to examine Air
Independent Propulsion (AIP)} but has not yet
found systems that are compatible with Australia's
operational profile. This is because of our enormous
transit distances, where the submarine might be
better off carrying more fuel and food rather than
a relatively cumbersome AIP system - though that
could change.
\end{tcolorbox}

These series of comments by RADM Sammut are consistent with other comments of his own and other government sources, and wholly inconsistent with a competent and sound engineering and design process.

It is inconceivable in an engineering project of this scale that after two years of work, and the conclusion October 2017 of the length, diameter and displacement of the submarine \parencite{sbs2017} that the type of propulsion is not yet known with certainty, and the inclusion (or not) of AIP is also not yet known with certainty.  As discussed in my senate testimony \ref{SENmorrison} either of these technologies will alter load (of batteries, fuel, and AIP reactants) of the order of hundreds of tonnes, in order to meet a fixed performance requirement.  Making such alterations within a hull that has its displacement already fixed could lead to catastrophic failure, essentially compromising the submarine's ability to sink or float.

Only three possibilities exist which can explain why such a statement has been made.

One possibility is that a firm decision actually already has been reached on the inclusion or exclusion of these technologies, and RADM Sammut is not willing to reveal it.  It should be noted here that the 'current thinking' wording might be interpreted as something of a retreat from earlier 'we are using' statements made in late October 2017 \parencite{sammut2017}.  The possibility that a pumpjet has already actually been ruled out must at least be considered.  In this scenario, it is almost certainly the case that the public and Parliament were misled on multiple occasions.

A second possibility is that RADM Sammut's comments are genuine and true, no firm decision about technologies except the exclusion of lithium and the size of the hull have been made, in which case the future submarine project could be headed towards catastrophic failure in exactly the same trajectory of the Super-Seasprite, by allowing massive changes to the included capabilities and technologies after the selection of a fixed platform of finite capacity.

The third possibility is that the performance specifications actually haven't been fixed either, and defence is planning on accepting whatever level of performance results from the combination of fuel, batteries, oxidants that fits in the defined hull-space in a neutrally buoyant and balanced arrangement driving whichever propulsor (propeller or pumpjet) is ultimately decided upon.  That such an approach to project management could be taken on a project like this is almost unthinkable from an engineering and risk-management perspective, not to mention cost-effectiveness.  However, in light of the evidence available, the third possibility may warrant quite serious consideration.

Such a possibility throws into sharp relieve some tensions between statements made by the government and Naval Group on multiple occasions.  RADM Sammut had previously stated that ``We did not specify that the boat should have or not have air-independent propulsion or lithium-ion batteries, rather that it met our requirements." \parencite{sammut2017}.  Weeks earlier, Jean-Michel Billig had cited said ``What is driving the choice of technology is basically the operational requirement" when challenged as to whether the pumpjet was actually mandated \parencite{lane2017}.  With both parties pleading ignorance of what technologies would ultimately be included, and both citing 'requirements', it does seem to imply that the requirements had not yet been settled.

If it was the case that the requirements had not been determined, then is likely that the Competitive Evaluation Process was essentially fraudulent.  If all the competitors weren't actually given the same performance requirements, which were the real requirements for the submarine which the government intended to procure, then it is extremely difficult to see how they could be `treated fairly' as RADM Sammut claimed \ref{SENhanson}. Moreover, credibility or conscience of the advisory panel who claimed they were should be seriously re-examined.  If the requirements have changed substantially since the CEP, then the CEP ultimately served no purpose.  A vast number of public statements made by politicians and officials regarding the significance of Australia's unique 'requirements', particularly for long range, should be critically re-examined.  It is likely that numerous statements made citing these requirements, and implying that they existed, were known, and were communicated to other parties including throughout and subsequent to the CEP, would be found to be misleading of the public, and the Parliament.

The explanation given regarding AIP adds weight to the bizarre notion that a back-to-front design process is indeed taking place.  The enormous transit distances cited have been well known for decades, and presumeably the requirement to meet these had been distilled into specific ranges and endurances long before the commencement of the CEP.  It's true that large amounts of fuel and food might consequently be required.  But the idea that we 'might be better off carrying more food and fuel' betrays either either a lack of confidence that the volumes which were presumeably propososed by all contenders alongside an AIP system were sufficient, or a casual aspiration to substantially enlarge the range and endurance of the submarine even further.  It would be relatively simple at the very early stages of concept design (prior to the diameter and legnth of the submarine being fixed) to incorporate both sufficient fuel and food for a specific overall range and endurance requirement, as well as an AIP system to enhance the dived endurance, without compromising the overall range and endurance.  That this was not done, and instead the dimensions of the hull were fixed, whilst Navy continued to search for a system which might fit within this pre-defined platform envelope does not remotely resemble a competent strategy to acquire a capability to meet a fixed requirement.

Worryingly, it adds confirmation that RADM Sammut might genuinely believe that the hull-diameter (and its resemblance of a pre-existing design) ought to be a pre-determined feature of a submarine design, to which the performance and capability of the submarine is subordinate.  Such a view, including held on the grounds of safety as suggested by RADM Sammut himself \ref{SENhanson}, is such a bizarre understanding of program risk and the heirachy of engineering challenges that it should call into question the competence of a program manager, and the technical advisors who he depends upon.

The competence of RADM Sammut's program management and the technical advisors on whom he depends is further marred by the comment about lithium-ion batteries being now firmly excluded from the design because the ``RAN does not want to be introducing a massive number of new technologies simultaneously".  This is telling because lead-acid batteries are known to have a relatively high internal resistance, which increases as they become hot, which occurs during levels of faster discharge.  This means that the efficiency of the batteries falls substantially at higher rates of discharge, reducing substantially the total amount of stored energy that they can usefully recover at very high discharge rates.  Whilst this is generally true for all batteries, Lithium-ion batteries are known to have substantially lower internal resistance, and consequently have dramatically improved performance (in terms efficiency) at higher rates of discharge when  compared to lead-acid batteries.  It is for this reason that in submarines, lithium-ion batteries are acknowledged as having the most significant improvement over lead-acid at high speeds \parencite{briggs2016}.  Given that pumpjets advantages are widely acknowledged to be at higher speeds, including implicitly by Naval Group\parencite{autret2016}, lithium-ion batteries should really be considered an enabling technology for pumpjets on diesel-electric submarines. (At least if the case is to be made that priorisation of that extremely high-speed sprint capability should be made above all other performance characteristics, which I have argued is the only technically defensible case to be made, whilst is an obscure and implausible case to be made in terms of useful operational capability \ref{SENmorrison})  Without lithium-ion batteries, the pumpjet efficiency will sap performance at low speeds, and at high speeds (where pumpjet efficiency improves, albeit not as high as propellers) the battery efficiency will hobble performance.  This effect, not accounted for in the modelling in reference \parencite{morrison2018}, will probably result in reducing the maximum endurance of a pumpjet-submarine in a sprint at 20kt from over 2 hours (as that modelling suggested) to well under one hour.  With lithium ion batteries, it is plausible that around 2 hours at this speed might be sustained.  The exclusion of lithium batteries \textit{because} of the pumpjet's inclusion further indicates that RADM Sammut remains, or is pretending to remain, completely unaware of such basic interactions between technologies.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=black!5!white,colframe=black!75!black, title= \textbf{Sept 2018 - Department of Defence, Report for the Senate Economic References Committee}]
\textit{After being requested to produce a response to my paper (Reference \parencite{morrison2018}) by the Senate Economics References Committee in a public hearing on 7 June 2018, Defence produces a six-page report including the following.  It should be noted that it was explicitly requested by Senator Carr that this report be produced \text{it} to the committee's final report from the inquiry the Future of Naval Ship Building Industry on the 27th of June.  Instead of the three weeks requested, Defence took about three months to deliver this report to the Senate Committee.  The full response is in }
\\
\\
\begin{itemize}
\item The Department maintains the view that this conclusion is incorrect in the case of submerged submarine for the following reasons:
\begin{enumerate}
\item The efficiency of a propulsor, whether it is a pumpjet or propeller, for a submerged submarine \textbf{is related to the total submerged resistance and is essentially constant over speed}, noting the submerged resistance of a submarine is dependent on the total skin friction and the shape of the submarine.
\item The efficiency of a propulsor, again whether it is a pump jet or propeller, for a vessel at the surface is related to the total underwater drag submerged resistance plus the wave making resistance, which means efficiency varies with speed for a surfaced vessel. A propulsor for a surface vessel will have been designed for a specific speed, with propulsor efficiency falling away at off-design speeds.
\item The efficiency of a well-designed submarine pumpjet will be similar to, or better than, a well-designed submarine propeller across the range of operating speeds.
\item Propulsors are designed for specific purposes. Pumpjets offer specific advantages for submarines, especially with respects to the acoustic signature.
\end{enumerate}
\item There are two primary components of resistance to forward motion for a surface ship, the underwater drag created by the hull form and the wave making resistance. The underwater drag is comprised mainly of frictional resistance and is a form of drag due to the viscosity of the water and the shape of the body. Also, when a vessel is on the surface it will generate surface waves. The wave-making resistance is the resistance caused by the generation of these surface waves and is a function of the speed.
\item \textbf{Deeply submerged submarines do not experience this wave-making resistance. This means that the propulsive efficiency for a submerged body will remain constant over the speed range; whereas the propulsive efficiency drops off for a surface ship as it moves away from the design speed.} This is independent of the type of propulsor. Therefore, the propulsive efficiency of a pumpjet designed for a submarine does not vary significantly with the speed of the boat.
\end{itemize}
...
\\
\\
An accelerating duct will generate forward thrust, especially at low speeds, hence its use on ducted propellers on surface ships operating at low speeds (tugs and trawlers). Decelerating ducts on the other hand reduces internal velocity, thus improving the cavitation performance. For a pumpjet, the duct and the stators contribute to improve the regularity of the wake around the aft part of the submarine . \textbf{For a submarine, an accelerating duct is attractive} as it generates additional forward thrust, further increasing efficiency.

\end{tcolorbox}

The assertion that the resistance of a submerged body is constant with speed is such an easily demonstrable falsehood (as discussed in more detail in the subsequent section) that one can quickly dismiss the core of their argument against my conclusion, which depends on this false assumption.  The competence or intention of the authors is further exacerbated by numerous additional remarks.  Perhaps noteworthy is the bizarre suggestion that an accelerating duct might be desirable for a submarine.  As was discussed in some length in my paper, accelerating ducts have the effect of reducing the static pressure of the water at the point that they reach the blades, which actually increases the amount of cavitation, and makes them very noisy.  They are never considered to be remotely desirable for military submarines.  It would appear that the author didn't even read my paper, nor any other relevant literature (for instance Reference \parencite{jackson1992}), or was simply intending to confuse and mislead an entirely non-technical audience.

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{kortvspump.png}
\caption{A diagram taken from \underline(Fundamentals of Submarine Concept  Design) by  CAPT Harry A. Jackson USN (Ret.) \parencite{jackson1992}, illustrating the comparison between an accelerating ducted propeller (Kort Nozzle) and a pumpjet (which is a decelerating ducted propeller).}
\label{fig:drag1.png}
\end{figure}

\newpage

\section{The drag resistance of underwater bodies}
It has been asserted a number of times, both in reference \parencite{kerr2018} and \parencite{defence2018} that it is only surface vessels who experience substantial changes in the drag they experience at different speeds, owing to the wave-making resistance that they encounter.

\begin{quotation}
``Since submerged submarines do not have to overcome wave resistance, the efficiency of a propulsor can be considered independent of speed..."\\ - \textit{Dr David Kershaw, Chief of Maritime DSTG} \parencite{kerr2018}
\label{resistancequotekershaw}
\end{quotation}

\begin{quotation}
``The efficiency of a propulsor, whether it is a pumpjet or propeller, for a submerged submarine is related to the total submerged resistance and is essentially constant over speed" \\ - \textit{Department of Defence} \parencite{defence2018}
\label{resistancequotedefence}
\end{quotation}

\begin{quotation}
``6. There are two primary components of resistance to forward motion for a surface ship, the underwater drag created by the hull form and the wave making resistance. The underwater drag is comprised mainly of frictional resistance and is a form of drag due to the viscosity of the water and the shape of the body. Also, when a vessel is on the surface it will generate surface waves. The wave-making resistance is the resistance caused by the generation of these surface waves and is a function of the speed. \\
7. Deeply submerged submarines do not experience this wave-making resistance. This means that the propulsive efficiency for a submerged body will remain constant over the speed range; whereas the propulsive efficiency drops off for a surface ship as it moves away from the design speed. This is independent of the type of propulsor. Therefore, the propulsive efficiency of a pumpjet designed for a submarine does not vary significantly with the speed of the boat."  \\ - \textit{Department of Defence} \parencite{defence2018}
\label{resistancequotedefence2}
\end{quotation}

These statements again use a seemingly relevant truth, in order to distract or obscure a subsequent falsehood.  It is indeed true that surface ships do encounter wavemaking resistance, and that this varies substantially as a function of speed, and submarines do not encounter this  type of resistance.  It is true that skin friction, (which is the shear component of the drag forces of any body that moves through a viscous fluid which act along its sides) as well as pressure resistance (the forces normal to the hull, which are non-cancelling on the nose and the tail) comprises essentially all of the drag that applies to a fully submerged submarine, as shown in Figure \ref{fig:drag4.png}.  However, to claim that these forces are constant with speed is to claim the most bizarre and obtuse denial of the most basic and well-established physics in this respect.  These forces scale with velocity squared, a fact which has been widely known and accepted since Froude conducted his skin-friction tests in the 1870s. \footnote{It should be noted that our understanding of skin friction has improved somewhat since the 1870s.  In particular, the distiction between laminar and turbulent flows and the transition between these two regimes as demonstrated by Osborne Reynolds in the 1880s was significant, since it was discovered that laminar flows exerted a lower skin friction, and it is the turbulent flows which scale with velocity squared.  However, whilst some effort is made to attmept to achieve some laminar flow over bow-mounted sensors in submarines at certain speeds, it is widely accepted that such a large vessel as a submarine is almost entirely enveloped by turbulent flows at essentially all operating speeds, and consequently the $V^2$ relationship has been essentially universally accepted as the dominant relationship for this purpose, including for the caluclation of scalar drag coefficients (which represent the impact of the shape and smoothness of the hull) to which much design effort and literature is devoted.}  The results of those tests are shown in Figure \ref{fig:dragfroude.png}. As such, speed plays the dominant role in considering the efficiency of propulsors on submerged bodies, since they also have their efficiency fall away at off-design conditions.

It is simply not credible that the Department of Defence can genuinely hold such a utterly  naive belief that is so clearly contradicted by known experience and the most basic and abundant literature on this topic.  Were it the case that the resistance faced by a fully-submerged body was ``essentially constant over speed" then a submarine or torpedo could reach any speed that it liked by applying the same force, and would have unlimited speed.   This would mean that even a diesel submarine would also have unlimited range by virtue of it's ability to reach any speed  that it liked.  This assertion  by defence is so utterly and blatantly ridiculous that there is no plausible way that it could be genuinely held to be a true by any one of even moderate technical competence.  It is difficult to even believe that it could be held with any confidence by even moderately intelligent adults of little or no technical background.   It is difficult to avoid reaching the conlcusion that the statement was given in a state of extreme contempt for the intelligence of the intended reader, as it constitutes without any doubt a deliberate lie and active attempt to mislead.

Almost every available source or text makes clear that the drag in water generally scales with velocity squared, including submerged vessels.

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag4.png}
\caption{A breakdown of the components of resistance taken from Reference \parencite{Mollard2011} does indeed reveal that deeply submerged submarines encounter no wave-making pressure, but still encounter frictional resistance (which scales with velocity squared).}
\label{fig:drag4.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{dragfroude.png}
\caption{A table showing the results of Froude's original experiments used to measure skin friction of various surfaces in the 1870s, taken from reference \parencite{lewis1988}.  Note that the exponent of velocity is determined consistently to be about 2. }
\label{fig:dragfroude.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag1.png}
\caption{Taken from the appendix of \underline(Fundamentals of Submarine Concept  Design) by  CAPT Harry A. Jackson USN (Ret.) \parencite{jackson1992}.  Note that the $V^2$ which indicates the relationship with velocity squared relationship}
\label{fig:drag1.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag3.png}
\caption{Taken from Research Paper commissioned by Department of Navy (US) titled \underline{Elements of the drag of underwater bodies} \parencite{granville1976}.  Note that the $U^2$ which indicates the relationship with velocity squared relationship}
\label{fig:drag3.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{drag2.png}
\caption{Taken from conference paper by Professor Volker Bertram titled \underline{Submarine Hull Design} \parencite{bertram2012}, same equation can also be found in published textbook by same author \parencite{bertram2000}.  Note that the $V^2$ which indicates the relationship with velocity squared relationship}
\label{fig:drag2.png}
\end{figure}

\section{The weight and size argument}
Both Dr David Kershaw, Chief of Maritime Division at DSTG\parencite{kerr2018}, as well as Mr Brent Clark\parencite{clark2018} have both made arugments to the effect of suggesting that it was the Australian submarine's very large size which is what made the pumpjet viable, where it might not be for smaller submarines.  I responded directly to Clark's comments arguing that the size necessitated by the extra fuel and batteries a pumpjet demanded could hardly be counted as an advantage \parencite{morrison2018b}.  However, Dr Kershaw's argument reveals a different as to why it was that size was important.  His comments make it clear that it is actually the large weight of the pumpjet \textit{itself} rather than the additional fuel and batteries it required, that was significant.

This argument implies that a smaller submarine couldn't adequately provide sufficient buoyancy or balance to support the heavy pumpjet. If the submarine couldn't provide such buoyancy or balance, the consequences would be catastophic failure, specifically the submarine would tend to either tail-stand, or sink.  It would be unthinkable for a competent manufacturer, conscious of the weight of the pumpjet and such risks, to neglect to account for such risks.  Weights and balances are known and predictable quantities, which could most easily be accounted for in the design of a new submarine.

However, Dr Kershaw's argument is most comprehensively debunked by the known fact that pumpjets have indeed been successfully retro-fitted to existing smaller diesel-electric submarines.  As one would expect, none of the catastrophic failures eventuated.  It is reported that the Russian Kilo Class submarine B-871 'Alrosa' has and does take part in exercises, and is remains in active service.  However, given that no other conventional submarines have since been designed with or converted to pumpjet propulsion, it is difficult to conclude that the benefits of a pumpjet on this type of submarine outweigh the costs.  It is trivial, however, to conclude that the weight of the pumpjet itself is not prohibitive for a smaller submarine, including  one smaller than the Collins Class, such as the Kilo.

Furthermore, the French also fitted a pumpjet to an Agosta 70 class submarine for a trial, as confirmed by Brent Clark in the Senate \ref{SENmorrison}. This submarine is even smaller than the Kilo, around 1500 tonnes, exactly the type of 'coastal' submarine that Brent Clark said a pumpjet wouldn't make sense for \parencite{clark2018}.  One would assume however that if the weight of the pumpjet itself was likely to cause the submarine to tail-stand or sink, such a trial would not have even been attempted.

Further physical arguments could be made as to how the weight of the pumpjet ought to scale with the size of submarine.  There are obvious reasons why the width/height (which must be equal, since the jet is cylindrical/conical) ought to scale in proportion to the diameter of the hull, in order to engage a proportionally larger water-column to produce the thrust necessary.  Similarly, the length  of the shround is likely to remain relatively fixed in proportion to the diameter for a given performance requirement. It is unlikely that the proportional thickness of the blades or shroud could be greatly diminished at larger scales without structural integrity and hydrodynamic qualities of the shroud and rotor blades being compromised.  All of these relationships strongly suggest that Dr Kershaw's implied suggestion that the weight of the pumpjet does not scale consistently with the weight of the submarine is completely unfounded, probably thoughtless, or otherwise  deliberately misleading.  However, given the clear-cut evidence that the type of failure that an over-weight propulsor would cause have been easily avoided on \textit{existing} smaller conventional submarines, such arguments wont be expanded in detail here.  It is abundantly clear that the arguments are entirely baseless, and demonstrably so.

It is difficult to imagine what sort of circumstance would lead both an acting CEO of the contracted design partner, who was aware of the experience of the French experiment with the Agosta, as well as a very senior defence scientist to make such demonstrably false claims in public.  In the case of Dr Kershaw, it can be seen that he deliberately and repeatedly (at the conclusion of his interview  \parencite{kerr2018}) drew attention to the size being significant.  His concluding remarks for the interview were that ``The size of the Future Submarine means a pumpjet is a real game changer for optimising the performance of the submarine in its intended role...". To deliberately mount an argument that is so simply ridiculed and dismissed as being so significant seems to be a bizarre choice for a supposedly competent scientist to make.  If stronger arguments were available, one would assume that they would be made instead. With over a million dollars already spent on modelling and experiments, one would assume that if the results of that research could confidently refute the conclusions my own research \parencite{morrison2018}, such a desperate and embarrassing statement would not need to be made.

In a clear parallel with the argument that the performance of propulsion systems on submerged bodies is independed of speed because they don't face wave-making resistance, this line of reasoning involves the presentation of a seemingly relevant, technical, yet understandable truth which is used to justify a completely blatant technical falsehood.  Even if Dr Kershaw is not familiar with this area of science and speaking outside his competence, it is difficult to see how such a pattern of argument occuring so consistently by accident.  For Dr Kershaw to genuinely believe in the arguments that he is making here is almost unthinkable.  For him to casually or accidentally make such careless arguments also seems implausibly reckless.  A more likely explanation is that the results of the research already conducted by the government broadly support the conclusions of my own research, and this is well known by Dr Kershaw, and  that these unfortunate statements are being made in a desparate attempt to deflect attention in order to avoid admitting as much.

\newpage

\section{The interaction between the hull and the propulsor}\label{hullpropulsor}

The argument that the 'hull needs to be tuned to pumpjet' and \textit{vice versa} has been used to dismiss critcism of the pumpjet repeatedly by RADM Sammut \parencite{sammut2017} \parencite{bergmann2018}, and also by Dr David Kershaw \parencite{kerr2018}. It has also been referred to in an answer to a Question on Notice given by Senator Ketter\ref{QON1}.  The answer to the Question on Notice, and also Dr Kershaw's comments comprise clearly the more detailed explanations, which  will be referred to in principal in this section.   However, the reference of RADM Sammut to this line of argument on multiple occasions, almost a year separated both before and after the publication of my paper\parencite{morrison2018} and my senate evidence \ref{SENmorrison} clearly indicates that it is a significant, perhaps the most significant line of defence of the  pumpjet in the program chief's mind.  The written response supplied by Defence to the Senate Economic References Committee also refers clearly to this argument \parencite{defence2018}.

As has already been noted, there are glaring contradictions from a project-management perspective which oppose this argument.  In particular, given that we are, according to RADM Sammut accept the pumpjet that was offered by France, and our submarine supposedly has a near identical hull diameter and  length  to the nuclear Barracuda, it is peculiar that any adjustment at all of either the aft section of the submarine, or the pumpjet itself, ought to be contemplated. The major advantage that the Naval Group proposal \textit{ought} to have is that those potentially complicated engineering challenges can be safely and entirely avoided.  Furthermore, the idea that the hull is what necessitates the tuning, as opposed to the speed profile, is almost incomprehensibly naive, as I pointed out in my evidence given to the Senate \ref{SENmorrison}.  It is noteworthy that no officials from Naval Group have ever alluded to this 'tuning' argument, or any adjustments to be made to the pumpjet or the hull on account of one another.  However, it seems pertinent here to give a reasonable discussion of the relevant physics nonetheless, in order that the public might fully assess the sort of argument being made by our most senior defence officials.

\subsection{The impact of the wake-field}

It is true that the hull does indeed change the flow of incoming water to a propulsor.  The area behind the vessel where the water is on being slightly dragged along with the ship (or the water is moving less quickly past the vessel, in the vessel's reference frame) is known as the wake.  Since any aft-mounted propulsor has to operate in the wake, or wake-field as it is often called, it is true that wake is of importance to the design of the propolsor.  What is not true, however, is that this is grounds for disregarding open-water efficiency, or that it was neglected from consideration in the preparation of my report.

In reality, the operation of a propulsor in the wake-field is an absolutely standard consideration for deriving propulsive efficiencies, as can be seen in any number of standard texts \parencite{Carlton2007}, \parencite{Mollard2011}, \parencite{MAN2017}, \parencite{lewis1988}. There is a ratio which is commonly discussed, relative rotative efficiency (denoted $\eta_R$), which has as it's precise purpose to describe the ratio between the efficiency of the propulsor in open water ($\eta_O$), and the efficiency of the propulsor in the wake of the hull ($\eta_B$).

It is widely acknowledged (for example in all of references \parencite{Carlton2007}, \parencite{Mollard2011}, \parencite{MAN2017}, \parencite{lewis1988}) that relative rotative efficiency generally lies around 1, and does not vary greatly.  For instance section 16.3.2, page 411 of reference \parencite{Mollard2011} says ``$\eta_R$ ranges typically from 0.95 to 1.05 and is often assumed to be unity for preliminary design purposes." \footnote{Observers may note that efficiencies are generally thought to range between 0 and 1.  The possibility that $\eta_R$ could be greater than one is opened up by the fact that it is actually a \textit{relative} efficiency, one efficiency divided by another.  Some propellers can actually work marginally better in the wake field than outside of it.}  The argument that the efficiency of pumpjets cannot be  assessed using their open-water efficiency is also blatantly false.  The 'open-water' condition is can be usefully thought of as the normal or 'ideal' efficiency of the propulsor, when the stream of incoming water is undisturbed, and consequently is still a very functional maximum efficiency of a system who's purpose is to smooth out the inconsistency in the flow.  No higher efficiency could be reasonably expected to be achieved through such  smoothing than would be acheived by beginning with a smooth stream in the first place.  Consequently, one can have considerable confidence that the extent to which this 'smoothing' effect might assist a pumpjet, relative to a propeller, is bounded by the extent to which it actually impedes a propellers performance.  As is amply demonstrated by abundant sources, this is a very small extent in all cases, and may actually be to the slight advantage of a propeller in some cases  where a desirable pre-swirl already exists.

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{roteff1.png}
\caption{One text's explanations of the variance between open-water efficiency and behind a hull from page 152 of \parencite{lewis1988}}
\label{fig:roteff1.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{roteff2.png}
\caption{One text's explanations of the variance between open-water efficiency and behind a hull from page 131 of \parencite{Carlton2007}}
\label{fig:roteff2.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{roteff3.png}
\caption{One text's desciption of the various types of efficiency from page 14 of \parencite{MAN2017}}
\label{fig:roteff3.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{roteff5.png}
\caption{One text's explanations of the variance between open-water efficiency and behind a hull from page 16 of \parencite{MAN2017}}
\label{fig:roteff5.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{roteff4.png}
\caption{One text's explanations of the variance between open-water efficiency and behind a hull from page 411 of \parencite{Mollard2011}}
\label{fig:roteff4.png}
\end{figure}

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{roteff6.png}
\caption{A paper from named \underline{Hydrodynamic Support in the Design of Submarines} also uses the same equations and relationships such as relative rotative efficiency to describe propulsion efficiency as that used for ships also uses \parencite{terwisga1988}}
\label{fig:roteff6.png}
\end{figure}

\subsection{Turbulence and Flow Separation Re-Capped}
It is noteworthy that absolutely no effort is made to engage in a discussion of turbulent or laminar in the boundary layer, nor does Defence make any mention of `flow separation' which are all important to the degradation of pumpjet efficiency at very low speeds.  A basic understanding of these concepts would be irreconcilable with the view that resistance underwater is constant with respect to speed.  In Reference \parencite{morrison2018} there is a substantial description  of these effects, particularly in Appendix 3 and Section 5, which have been apparently ignored by the Department of Defence.  Some basic re-capping is suitable here.

When any solid surface moves through a fluid (or \textit{vice versa}) there is always some boundary adjacent to surface, where the the fluid is actually non-stationary, and moves along partly with the solid.  This is intuitively named the `boundary layer'.  There are two different types of boundary layer.  A \textit{laminar} boundary layer has all the layers of the fluid in the boundary moving in parallel, with the speed of each layer reducing further away from the surface.  Such  a type of flow tends to only happen over very small distances, or at very low speeds, or in very viscous fluids.  For the majority of the surface consideration of a submarine, the boundary layer is actually turbulent, which means that water in the boundary layer is tumbling and turning in a confused and unpredictable pattern.  Whilst \textit{on average} the boundary layer moves closer to the surface speed closer to it, and slower further away from it, at any local point on a tiny lengthscale the movement of the water will be chaotic, and altogether very non-parallel.  Turbulent boundary layers tend to be much thicker, and become even thicker at higher velocities, resulting in larger amounts of energy being sapped producing the faster, more intense chaotic swirling motions that comprise turblent flows.

This distinction is important in understanding the limitations of a pumpjets claim to make the incoming stream of water less inconsistent, or more regular.  The duct intake, and any stator blades in advance of the rotor can be effective in making the flow more homogenous, but in doing so they simultaneously make the flow more turbulent, since the necessarily involve passing the water over more surfaces, rather than less. This can be advantageous to the extent that very large-scale and intense vortices might be broken down into smaller-scale vortices, which appears more homogenous when assessed on the length-scale relevant to the blades of the rotors.  This could indeed have the effect of reducing the likelihood of very large pressure fluctuations impacting the blades, reducing the probability of cavitation occurring or invoking some kind of resonant beats within the propulsor.  \footnote{It is this effect that is likely referred to in comments made by Professor Joubert in a paper commissioned by DSTO in 2004 \parencite{joubert2004}.  However, with competently designed appendages and fairings, the kind of unsteady vortex shedding which would see such large-scale eddies enter the propulsor should be safely avoided.  Substantial effort has been devoted to developing the smoothest possible wake-field for submarines for this purpose. \parencite{toxopeus2014}}

However, it must be noted that the in making the flow more homogenous, the flow is also made more homogenously turbulent, and the total level of energy which is wasted to  small-scale chaotic turbulent movements necessarily increases, as it necessarily must to with any additional surface. The circumstances where a pumpjet is able to attain higher efficiency are those where the total amount of additional productive energy (thrust, produced by front to back acceleration) added to the water column dramatically exceeds the losses incurred by the additional turbulence (round and round acceleration).  This occurs at high speeds.

It should be noted that a pumpjet's duct also has the effect of decelerating the incoming flow, which increases its static pressure.\footnote{In fact, a pumpjet is by definition a decelerating ducted propeller, contrary to claim that they are so different as to be incomparable made \ref{QON1}.} Whilst this has the desirable effect of further reducing the likelyhood of cavitation occurring, it also creates an `adverse pressure gradient' in the intake.  This means that the water must move towards an area of higher pressure, agains the direction that the pressure would naturally push it.  Whilst at high speeds the water has enough momentum to easily overcome this resistance, at very low speeds there isn't enough energy in the flow for the entire water column to move consistently in this direction, and part of the flow simply turns around and moves in the opposite direction, in a phenomenon known as `flow separation'.  This occurs against a surface (in this case just inside the pumpjet intake), where the boundary layer has to move more slowly than the remainder of the stream, and in  this case it is the first part to reverse direction.  Whilst the entirety of the boundary layer, (or indeed the entire flow) may actually be turbulent, with many very small-scale vortices, this phenomenon creates new large-scale macroscopic vortices, which could form, grow and shed in an unsteady creating exactly the same risk of perodic shedding (i.e. beats) impacting the rotor that a poorly designed mast or appendage might have.\footnote{A cylindrical mast, for instance, could set up a Karman Vortex Street at certain speeds.} Whilst it might be presumed the shedding of vortices created by flow separation into the pumpjet at low speeds has been adequately managed or mitigated to the extent that it does not risk the creation of cavitation on the blades, it would most certainly lead to the very substantial loss of efficiency through creation of these vortices. Since open propellers do not have a duct which creates an adverse pressure gradient, they do not face this problem at low speeds.  In fact, at very low speeds more substantial parts of the blade may actually have laminar flow over them, even further enhancing efficiency.

\begin{figure}[h]

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{HiFlowSmall.png}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{LoFlowSmall.png}
\caption{}
\end{subfigure}

\caption{(a) At high speeds, the intake of a pumpjet slows the waterflow and raises its pressure, and it subsequently accelerates through the contracting nozzle to form an efficient jet (b) At low speeds, the pressure gradient created by the duct causes some of the waterflow to stop and reverse.  This `flow separation' wastes energy spinning water in circles, and dramatically reduces the system's efficiency.}
\label{fig:HiLoSmall.png}
\end{figure}

In clear contrast to the trivial changes that the relative rotative efficiency $\eta_R$ incurs,  the open-water efficiency $\eta_O$ of a propulsor varies between 0\% and 75\%, depending crucially on the advance ratio $J$, which is highly speed dependent due to the changing resistance of the vessel at different speeds.  In \parencite{morrison2018} it was shown that it was most probable a pumpjet's open-water efficiency could be around half that of a propeller at speeds at or below 5kt.  It is consequently vastly more important to give detailed consideration to early in the design and selection of a propulsor, particularly for a diesel-electric submarine.  Unless defence is proposing that a pumpjet can have a relative rotative efficiency of more than ~\%200, there is no feasibly means that this line of argument can sustain their dismissal of my conclusion.

A particularly desparate line of argument might lead one to argue that such arguments do not apply to submarines propulsors due to their great level of depth and the unique shape of their hull. This would not be true.  Owing to their generally cylindical shape and generally conical afterbody, submarines tend to have wake fields are much less complex and intense in the degree of their variation that surface ships. This is because ships bodies also have wave-making effects involved as well as a hull that tapers above the propulosor, creating an uneven effect with different heights.  In contrast, wake fields of propellers have a relatively even gradient in wake fraction (the intensity of waterspeed difference due to the hull) in a radial direction, with only the control surfaces ahead of the propulsor causing substantial variations.  Furthermore, at very low speeds, where pumpjets were shown to have the lowest efficiecies, the effect of the wake field would be the least compared to the open-water case, since the size of the boundary layer about the aft of the hull would be substantially diminished.  On absolutely all accounts, the interaction between the hull of a submarine and propulsor at low speeds cannot be credibly argued to have any impact which would overturn the conclusion reached about pumpjets having clearly and substantially inferior efficiency at low speed, as argued in \parencite{morrison2018}.

As such, the reason that my earlier paper did not explicitly discuss relative rotative efficiency and the impact of the wake field, this was due to it being so obviously and demonstrably of no bearing whatsoever on my conclusions, and I had assume that as much would immediately clear to any technically competent reviewer.  The use of this argument displays a degree of ignorance of the physics, as clearly described in numerous and publically available sources, that it is difficult to account for by incompetence alone.  A substantial measure of contempt for public intelligence must be required in order to appeal to such trivially simple statements of truth in order to support conclusions which can be so trivially demonstrated to be completely false.

Imagine that you came to buy a coffee from a your favourite cafe, only to find that they weren't selling any coffee that day, and wouldn't be selling any more indefinitely.  Their coffee maching had been removed, and they were only selling cold-drip, coffee, stored in little bottles in the fridge. Overcoming initial shock (since you weren't aware of any other cafes who would make such a bizarre decision as to not sell espresso-coffee as a substantial part of their offering), you made some tentative inquiries as to why.

At first, you're told vaguely that it was to do with health and safety.  This seemed a little odd, since there obvious benefits to everyone's enjoyment and presumeably the business, which you point out to the cafe owner, and you ask how the health and safety risk could really be so large as to outweigh that.  With just a little annoyance at having to explain, the cafe owner briefly says something about temperature being the problem.  You look at the empty spot on the bench where the coffee machine used to be, and speculate that perhaps there really was some serious fault with the machine, maybe even a flaw in its design requireing it to be recalled.  Perhaps it was running uncontrollably hot, and might have risked melting componentry, creating electrocution risk, or even a risk of explosion if the boiler rose to an elevated pressure.  That would explain the safety.  Or perhaps there was some emerging science about carcinogens being produced in the high-temperature extraction of espresso, which cold-drip might presumeably avoid?  In anycase, it sounded like it could be very technical, and you decided to trust your very experienced and seemingly intelligent barista and forget about your lingering doubts as to why no other cafes were following suit.

Then later you find a notice posted up at the front of the cafe, explaining fully the rationale behind the abandmonment of the coffee machine, and the connection with temperature. In it, it is explained that all espresso-based coffees were hot, those with milk around 65 or 70 degrees, and long blacks or fresh espresso potentially above 90 degrees.  If consumed quickly or accidentally spilled such drinks could potentially cause burns.  Therefore in the interest of health and safety no more hot drinks would be served, and cold-drip would be the only coffee available.

You read the statement again, not quite believing that a hazard so obviously known and successfully navigated by millions of coffee drinkers every day, could be the cause of such a decision.  Althought it's true the owner didn't lie about it being a health and safety concern, and can't dispute the facts about the coffee temperatures being above 65 degrees, you still conclude that owner is either insane, or trying to disguise some other reason for abandoning the espresso coffee.

\printbibliography

\newpage

\begin{appendices}

\section{Extracts from public statements}

\subsection{Hansard Extract, Senate Estimates - 28th May 2012 \cite{SE2012moffit}}\label{SENMoffit}

\textbf{Senator JOHNSTON:} There were announcements made about these various studies. You have explained propulsion and energy storage. I get that. Signatures and stealth performance is self evident. You have touched on combat systems and the hydrodynamics of the propellers. But pump jets is the one that has the big question mark over it. Why are we looking at pump jets?

\textbf{Rear Adm. Moffitt:}  Why are we looking at it?

\textbf{Senator JOHNSTON:} Why are we looking at pump jets on a conventionally powered submarine?

\textbf{Rear Adm. Moffitt:}  We are not specifically ruling in or out anything at the moment. Pump jets are in use in submarines and we will-

\textbf{Senator JOHNSTON:} Mainly nuclear submarines with a speed far and away beyond 25 knots.

\textbf{Rear Adm. Moffitt:}  That is true, as I understand it. The issue is that we need to understand that, because we have not got much knowledge of it, before we rule it in or out. There are many new technologies, including pump jets, AIP and lithium batteries, that have emerged since we executed the Collins program about which we know relatively little and wish to become more educated about than we are today before we recommend to government that they be included or discounted. We just do not know enough to give a sensible answer to government at the moment. I am not saying that pump jets are in. Nor am I saying that they are out. We just do not know enough yet.

\textbf{Senator JOHNSTON:} What is our methodology to get that information? Who are we asking? How much is all that costing?

\textbf{Rear Adm. Moffitt:}  We will be relying fairly heavily on DSTO to help us with that. We will also be fairly keen to get some advice from the US Navy on that as well.

\textbf{Senator JOHNSTON:} I am told that there are no pump jets on any conventional class submarines in the world.

\textbf{Rear Adm. Moffitt:}  That is quite right. It might be a very easy task for us to write it off. I do not know, because I have not done-

\textbf{Senator JOHNSTON:} The French and the Russians have tried. No-one has used them. They are not being used on anything other than nuclear powered submarines-particularly the Americans.

\textbf{Rear Adm. Moffitt:}  If it does not look like it makes sense, I will rule it out. But I cannot give you the answer to that yet because we have not done the work.

\textbf{Senator JOHNSTON:} Who had the idea to look at pump jets? Where did that come from? Did it come from capability group saying that we had better had a look at that?

\textbf{Rear Adm. Moffitt:}  I do not know the answer to that. I doubt that there is a clear answer. It is a submarine technology that we think that we need to know more about than we do today. I do not anticipate spending a substantial amount of money to get the understanding that I am looking for-I may not have to spend anything at all.

\newpage
\subsection{Hansard Extract, Senate Estimates - 25th October 2017 \cite{SEN2017hanson}}\label{SENhanson}

\textbf{Senator Hanson:}   I have a question about the Shortfin Barracuda subs. Just to get it on the record, is it correct that the cost of the contract is \$50 billion?

\textbf{Rear Adm. Sammut:}   The cost of the program is \$50 billion in acquisition. That is not the cost of the contract per se; that is the costs that will be required to establish the submarine capability in full, which extends to not only the acquisition of the submarines but the
establishment of the facilities necessary to operate the submarine; the establishment of a shipyard to be able to build those submarines; wharfs; training centres; test and integration facilities for the combat system; the acquisition of the combat system; and the establishment
of a sovereign capability in Australia as well.

\textbf{Senator Hanson:}   Is that for a total of 12 submarines?

\textbf{Rear Adm. Sammut:}   That's for the submarine capability, which comprises, as part of
that, 12 submarines.

\textbf{Senator Hanson:}   When was the contract signed?

\textbf{Rear Adm. Sammut:}   There will be several contracts signed for the delivery of the Future
Submarine. The first contract was signed in September last year. That was known as the
design and mobilisation contract, under which we have commenced design activities and
activities in terms of mobilising both Naval Group and Lockheed Martin, along with the
Commonwealth, to commence that design process, which will continue over the next several
years. By way of example, we have commenced the design of the submarine itself. We've also
established a resident project office in Cherbourg. We have commenced basing our people in
France. We've started the development of our engineering plans, the establishment of an
integrated master schedule and so on.

\textbf{Senator Hanson:}   At what time prior to that did you do investigations around other
submarines-how many months prior to signing that contract? How much time and effort did
the Defence Force put into looking at other subs that were available prior to the Shortfin
Barracuda?

\textbf{Rear Adm. Sammut:}   We commenced the competitive evaluation process in 2015 and
work on other options had been proceeding from about 2011. So there was about four years of
work looking at various options that included military off the shelf, modified military off the
shelf, and evolved Collins.

\textbf{Senator Hanson:}   With regard to the Shortfin Barracuda, you have signed a contract
with Naval Group, which is a French firm. Did you specifically go looking for a diesel
submarine? Or what were you looking for?

\textbf{Rear Adm. Sammut:}   The competitive evaluation process was for a conventionally
powered submarine, a diesel-electric submarine, not any other form of submarine.
\textbf{Senator Hanson:}   Was it based on a pump jet sub?

\textbf{Rear Adm. Sammut:}   What we were looking for was based on the best solution that the
three international participants within the competitive evaluation process could offer to meet
Australia's capability requirements. We did not specify to any of the participants the form of
solution it should take, the form of the submarine. It was up to those international participants
to present their design and, more importantly, their design capabilities to meet Australia's
capability requirements for its Future Submarine.

\textbf{Senator Hanson:}   You didn't want a nuclear sub, that was completely out of the
equation?

\textbf{Rear Adm. Sammut:}   That's correct, a nuclear submarine was not in consideration. The
government policy was that we were to acquire conventionally powered submarines, the
Future Submarine.

\textbf{Senator Hanson:}   So you were looking at either a pump jet or a diesel?

\textbf{Rear Adm. Sammut:}   I think you're confusing, Senator, a pump jet as being a nuclearpowered
submarine, versus what is in fact a conventionally powered submarine. A pump jet is
a form of propulsion. It's not a form of power for the submarine. It is a form of propulsion that
fits to the back of the submarine and, rather than taking the form of a conventional propeller,
it takes the form of a shrouded rotor with a set of stators in there that, in operation, creates
thrust through the water.

\textbf{Senator Hanson:}   All right. How are the Shortfin Barracuda run-on what?

\textbf{Rear Adm. Sammut:}   I think the point that we're trying to make here is that the proposal
offered by France included a pump jet.

\textbf{Senator Hanson:}   So it is pump jet? So you're going pump jet?

\textbf{Rear Adm. Sammut:}   The proposal-

\textbf{Senator Hanson:}   This is why you've chosen them as the design-to go with the pump
jet. So it's not propellers, is it?

\textbf{Rear Adm. Sammut:}   The proposal offered by France included a pump jet, and that's the
proposal that we've accepted. A pump jet remains on offer, and that's what we intend to use
for the Future Submarine.

\textbf{Senator Hanson:}   And that's why you chose France's Naval Group to-

\textbf{Rear Adm. Sammut:}   No, that is not, alone, why we chose France. There are a number of
things. I am on record as providing testimony at the Senate of the number of things we
considered in the selection of Naval Group as our international partner. There were a range of
factors, and they were all over the capability of the international partner to deliver a regionally
superior submarine and one which Australia would have the sovereign capacity to operate and
sustain throughout its life. There were a number of criteria in mind.
We looked at the capabilities. That includes the platform capabilities. We looked at the
ability of the international partners to work with the combat system integrator of our choice.
Those things in particular pertain to the capability of the boat. We looked at rough order of
magnitude costs and schedules to ensure the participants understood the scale of the program
that we were embarking upon and could give us insights into the manner in which they would
manage costs and schedules throughout the program. We looked at matters concerning
program management. We looked at the design and the way that safety is incorporated into
the design. It's important that that was understood from the outset because, whilst the
international partner would be designing the submarine, it would be the Commonwealth that
would be responsible for certifying the submarine as safe and fit for operations, and we
needed to make sure that the certifying process would match the design philosophy.
We looked at methods of sustainment. We looked at the manner in which each of the
participants would support the establishment of a sovereign sustainment capability in
Australia. Of course, that was a key lesson from the Collins program. Even though we built
the submarines in Australia, it took us some time to establish the sovereign capacity to sustain
those boats. Indeed, having done that now is the reason why we are able to deliver beyond
international benchmark availability. We also looked into the way in which data would be
transferred to Australia, the manner in which intellectual property would also be managed in
the program. We looked at the ability of the participants to manage risks. Again, throughout
this complex program, that would be a very important factor. I would say that we didn't just
pick a submarine with a pump jet; we picked an international partner that could meet a range
of requirements to achieve not only a regionally superior submarine but the sovereign
capacity to operate and sustain that submarine and to deliver that within the time frame that
would be necessary.
I'd like to point out that that process was peer reviewed and it was peer reviewed by an
expert advisory panel. The expert advisory panel found that the process was sound and that
the participants had been treated fairly. I might add that the expert advisory panel also met
with each of the participants on several occasions to ensure that they were satisfied that the
process was being run competently and soundly. I'd also say that the results of our work were
peer reviewed by former program managers from the United States, who made no
recommendations to change the course and the manner in which we were conducting the
competitive evaluation process and indeed came to the conclusion that the manner in which it
was being conducted would indeed lead to appropriate recommendations on the selection of
an international partner.

\textbf{Senator Hanson:}   That is correct: you did get advice from the United States, from US
Navy secretary Don Winter, retired US Navy Vice Admiral Paul Sullivan and Rear Admiral
Thomas Eccles. They were actually on this based on sound experience. But the fact is that
America had not launched a diesel submarine for 55 years.

\textbf{Rear Adm. Sammut:}   This was about a submarine program. These are people with a lot of
experience in submarine design. Might I also add that we have a lot of experience in the
operation of diesel-electric submarines ourselves. We understand many of the factors around
battery technology. I'd just like to add-

\textbf{Senator Hanson:}   Sorry, I have limited time with this. I just want to ask you-

\textbf{Senator GALLACHER:} Surely courtesy on both sides is required here.

\textbf{CHAIR:} Senator Hanson has the call.

\textbf{Senator Hanson:}   I agree, but I need to ask these questions. I just have to say: you're
actually going to put pump jets into a sub. Is it true that pump-jet submarines can stay
underwater for only 20 minutes?

\textbf{Rear Adm. Sammut:}   No, I do not know what you're referring to about pump-jet
submarines remaining underwater for 20 minutes. Can I be clear again: a pump jet is a form
of propeller. It has no bearing on how deep a submarine can go or how long it can stay
underwater. They are matters that pertain to the platform design of the submarine.

\textbf{Senator Hanson:}   What about a diesel sub? We are going diesel.

\textbf{Rear Adm. Sammut:}   We are going to buy diesel-electric submarines, and I can assure
you they are going to stay underwater for much longer than 20 minutes.

\textbf{Senator Hanson:}   Three days? What's the period of time?

\textbf{Rear Adm. Sammut:}   I am not going to reveal what our requirements are for submerged
endurance. That is classified.

\textbf{Senator Hanson:}   I don't think it's classified.

CHAIR: You don't get to make the call on what's classified and unclassified in this
conversation.

\textbf{Senator Hanson:}   So you can't disclose knowledge of how long a diesel sub can stay
underwater compared to a nuclear sub?

Mr Johnson: A diesel-electric submarine can stay underwater depending on the way it's
designed. That refers to matters about the size of the battery, the manner in which the hotel
load, which is the power that is consumed by the submarine when it's operating, is used and
so forth. There are a number of factors that go into the period of time that a diesel-electric
submarine can remain submerged.

\textbf{Senator Hanson:}   This Barracuda-there is still one on the slipways in France that
hasn't even been tested. We're basing the design on that which has not been tested. Do you
believe that is in our interest?

\textbf{Rear Adm. Sammut:}   We have selected Naval Group as competent designers of dieselelectric
submarines and indeed ocean-going submarines. We made that assessment on the
basis of the competitive evaluation process, which, as I've endeavoured to explain, was a very
thorough process. We have no reason to believe that France is not capable of designing a
submarine to meet our requirements.

\textbf{Senator Hanson:}   You don't believe we may be making the same mistakes as we did
with the Collins class when that was first designed, which cost us billions of dollars?

\textbf{Rear Adm. Sammut:}   No, I don't, because we've learned many lessons from the Collins
program and implemented those as we've gone forward with the Future Submarine.

\textbf{Senator Hanson:}   Is the Barracuda designed as a nuclear sub-that is, basically is a
nuclear sub?

\textbf{Rear Adm. Sammut:}   The French Barracuda, the submarine that they will be employing
in their navy, is a nuclear powered submarine. The submarine that we're working on with
Naval Group is a diesel-electric submarine.

\textbf{Senator Hanson:}   But it is still the Shortfin Barracuda, which is a nuclear sub, so we're
taking a nuclear sub and we're going to make it a diesel sub.

\textbf{Rear Adm. Sammut:}   No, we are designing a submarine that will have a certain lineage in
the work that France has done in designing both the Barracuda and conventional submarines.
That's quite natural in any submarine design and build process. If you looked at any nation
designing and building a submarine, they will always refer back to the lineage of designs
they've done. What we have the benefit of doing here with France is leveraging the design
work that they've done in the Barracuda and a number of conventional submarines that
they've designed, built and delivered to other navies, and using that as the basis for the design
of the Future Submarine, which I said, is a conventionally-powered submarine. You cannot
convert a nuclear submarine to conventional power. That has often been stated in the media,
and I understand most recently in the Insight Economics report. The effort is not to convert an
existing boat into a diesel-electric submarine; we're designing a submarine using many of the
learnings and references that come out of the Barracuda design. A good example there would
be something like hull diameter. If you're going to change the diameter of the hull of a
submarine, you're into a whole new set of calculations about the structural integrity and
strength of the submarine. If we can utilise the work that France has already done in having a
hull diameter that's very similar to Barracuda, we can leverage all of that design work, which
has not only gone into the Barracuda but also has been tested by France in their methods of
design and certification to get to this point.

\newpage
\subsection{Hansard Extract, Senate Estimates - 15th December 2017 \cite{SEN2017johnson}}\label{SENjohnson}

\textbf{Senator KITCHING:} Are the future subs going to have pump jet propulsion?

\textbf{Mr Johnson:} Yes, they will. Given you're interested, I might take a moment or two to explain.

\textbf{Senator KITCHING:} Yes, sure. Thank you.

\textbf{Mr Johnson:} The method of designing for either a propeller system or a propulser system, a pump jet, is very similar. It starts with key parameters: how fast do we want to go, how big, what's the weight and diameter of the vessel that we're going to power, is it a surface ship or a submarine-although the calculations are really different-and then what's the shape of that hull? All of those things then establish the particular characteristics of a propeller or a propulser pump jet design. The submarine that we're building is well suited, and the requirements that we have for performance are stringent that the pump jet technology is the optimum approach. Since it's a new design-contrary to misunderstandings in the press, this is a new design-we're in a good position to do those calculations for the pump jet from day one and to balance the submarine's designs so that it performs very well with that pump jet technology. That's been our plan since the selection of what was then DCNS and is now Naval Group France.

.
.
.

\textbf{Senator PATRICK:} I go back to pump jets, if that's possible. Maybe the Chief Defence Scientist could come to the table if he's about. He might be able to help.

\textbf{Senator Payne:} Dr Zelinsky will be here somewhere.

\textbf{Senator PATRICK:} Mr Johnson, there's been a lot of talk about pump jets. I've now tabled a document which has come from the United States, from Connecticut. You'd be very familiar with the organisation there. That basically describes the characteristics of pump jets, their pros and cons, as any piece of engineering equipment has. The document-and there's also another document that I've taken from an ASPI site-shows the efficiency characteristics of a pump jet. Both the US document-now declassified-and the document from the ASPI website suggest that, at low speed, pump jets are relatively inefficient compared to, say, a propeller. That may well be the reason why we see a pump jet on a Mark 48 torpedo but generally not on merchant ships and so forth. Would you care to comment on that?

\textbf{Mr Johnson:} As we said just a little earlier, the design process-whether it is a propeller or propulsion-is very much keyed to the criteria that you look for. The first thing I said was speed. No matter what kind of propeller or propulsion pump jet you design, it has to be optimised for a particular speed. There are classes of surface ships that look to optimise cross-ocean transit. Physics usually keeps you around 17 to 18 knots, and then there is the cost of maintenance and the cost of repair and the fact that there's no particular stealth requirement. Typically, you will find that those ships will use an open-tip propeller. However, other ships, cruise ships in particular, use an electric shrouded pod, which is a variation of a pump-jet technology. Of course, the technology gets changed over time, so the older a document is, the more it's dated to that point in time.

\textbf{Air Chief Marshal Binskin:} Can I just confirm-we've got two documents and I'm trying to make sure we've baselined ourselves here. We've got one document, which is the 1963 document.

\textbf{Senator PATRICK:} That's correct.

\textbf{Air Chief Marshal Binskin:} We haven't got the other document yet.

\textbf{Senator PATRICK:} It's just a diagram.

\textbf{Air Chief Marshal Binskin:} Thanks for that.

\textbf{Senator Payne:} I'm sure the officials can speak in the general, Senator, in response to your questions, but, if your questions go to the detail of that original larger document, you can hardly expect them to speak to the specifics of it.

\textbf{Senator PATRICK:} Sure.

\textbf{Mr Johnson:} We were talking about optimising for the speed and the characteristics of the pump jet. The decision-making process for submarines is a little more complicated because there's an acoustic performance aspect as well as the pure efficiency of the blade-propulsor-pump jet arrangement. Without going into details, it's a decision that we make with Navy, but I'm confident, having done this a few times already, that we are going to find this pump-jet approach to be the optimum for our particular circumstances for this particular submarine.

\textbf{Senator PATRICK:} The general summary of that larger document is that pump jets are most efficient at high speed. That's where you get the benefit because there's the cost of the propulsor and the weight of the propulsor. The shroud, for example, is great for reducing cavitation at high speed, which is typically what a nuclear powered submarine might do in transiting from one place to another. Indeed, to my knowledge, the only submarines that are fitted with pump jets, other than one Agosta 70 French submarine it was trialled on and one Russian submarine, are nuclear powered submarines. So the data in the public domain suggests that the advantages of the pump jet occur at high speed. The diagram that you're looking at shows efficiency at low speed, and that's where a submarine spends most of its time in an operational area. A question was asked by Senator Xenophon back in October about this efficiency problem. The answer came back that 'a submarine propulsor can provide higher propulsive efficiency than a propeller across the speed range of a submarine, including the patrol speed'. I don't want you to go to the technical details or anything that's classified, but are you saying about the diagram that I've presented, which is in the public domain, that there's now new technology that would be contrary to that?

\textbf{Air Chief Marshal Binskin:} No. I think what Mr Johnson's saying is that all aspects of our concept of operations will be factored into the submarine design in order for it to be best able to conduct the operations we need. So I don't think we should be talking in any more detail than that here in an open forum.

\textbf{Senator PATRICK:} Let me tell you my concern.

\textbf{Air Chief Marshal Binskin:} I think I understand your concern.

\textbf{Senator PATRICK:} Actually, it's a concern that relates to risk in some sense. When you put this really heavy pump jet at the back end of a submarine and then at some later stage you find that in actual fact the pump jet doesn't perform as you might have expected, noting what is in the public domain, to change that out, noting it's a big, heavy thing at the extremities of the submarine, will have significant impact on the design. So my question to you is: are you relying on data that is based on a real in-situ propulsor somewhere or are you relying on modelling data to make this decision?

\textbf{Mr Johnson:} The decision to use a pump jet was carefully made. It includes a wider understanding of their use in other applications-leaving that open. We have already accessed a relatively large scale model, so its physical data, and we will have the advantage of slightly different but very similar arrangements when the French attack submarine goes to sea. The risk that you note is certainly a valid concern, but the information that we have is across a wide spectrum of available sources, some not available to you, and the fact is that we will have this very early full-size, full-scale performance data of a different propulser, but nevertheless one from the same team in a very similar hull shape. At this point, the Commonwealth should be very comfortable with the calculated decision that we're making.

\textbf{Senator PATRICK:} Thank you very much, and I will come back to you briefly on that. Back in 2014, Senator Johnston was making inquiries as to the way in which the Commonwealth was spending \$214 million of de-risking money for their submarines. In amongst that money was a study on pump jets. I will refer you to the Hansard of 28 May 2012, where it makes it clear that DSTO were doing some studies on pump jets. Are you aware of those studies and the outcomes of those studies?

\textbf{Dr Zelinsky:} I'm aware that we have been studying propulsers, propellers, various forms of propulsion for submarines, but I cannot, at this stage, give you the outcomes of that study from five years ago. I'd have to take that on notice. Subsequently, as we have gone through the competitive evaluation process, we have provided work and studies to that. I am sure our work has matured and our work does underpin the work of CASG.

\textbf{Mr Gillis:} This is a 1963 document, which took a number of years to be declassified to this point. As I think most of the people here know, submarine technology and what we decide, and how and why we decide it, is, by its nature, classified. We want to be really careful that we don't go into a point where the money and great effort that's been undertaken by our Defence Science and Technology Group, the work we've done with our partners, is then put on any sort of public forum for debating.

\textbf{Senator PATRICK:} I wasn't after the detail.

\textbf{Mr Gillis:} The issue we have is that we have some of the best scientists in the world; we have some of the most experienced submarine program managers in the world working at it. We partner around the world to ensure we get the best possible outcome. As the head of procurement, I trust that those people are making good, sound decisions not based on very dated data; they are based on very up-to-date data. And as you saw, it's a \$200 million-plus investment by the government, and one of the aspects was propulsion. I think you should be comfortable that we have invested the money. The issue is I have no intention of putting that sort of information in the public domain.

\textbf{Senator PATRICK:} Sure. All I'm after is just some indication from the chief scientist, noting there's a diagram on the table, as to whether or not he's comfortable. That is the question. I'm not asking him to go to the detail.

\textbf{CHAIR:} So is your question essentially: is the chief scientist comfortable that the science backing up the decision of Defence is sound?

\textbf{Senator PATRICK:} Yes.

\textbf{Dr Zelinsky:} Yes.

\textbf{CHAIR:} Thank you.

\textbf{Senator PATRICK:} Okay. I'll go back to you, Mr Johnson. Noting that you're saying that this is based on something that hasn't been to sea yet-it will go to sea on the Barracuda; presumably that's the submarine you're referring to-there's still a risk that's sitting there. As the guardian of the taxpayer, I know where you sit. I have a responsibility from the taxpayers' perspective to ask: do you have a fallback position if the data that comes from the French trials, or other trials that you might do, presents you with data that's inconsistent with the operational profile that you intend for the submarine?

\textbf{Mr Johnson:} Yes. We will understand the performance data early enough in the design that we can accommodate whatever is required. And then, just to reassure you, it's really going to be an adjustment to the design of the pump jet, not a decision to leave it. Without going into too much detail, the curve that you're referring to is about the efficiency of the propeller, not about its acoustic performance. As you know, there's nothing free in physics, so those are different-shaped curves. What we seek to do, as Chief of the Defence Force said, is to make the right balance of multiple dimensions of this now.

\textbf{CHAIR:} Senator, did you want a private briefing to go into the detail?

\textbf{Senator PATRICK:} No, I'm satisfied with that; I'm satisfied with the detail.

\textbf{Mr Johnson:} One other reassurance: we will run scaled model testing, and that will start relatively soon. It's a standard approach. It's been used by the French navy and the US navy for a long time. We're going to march right down the path, and we will do it with the chief scientist.

\newpage
\subsection{Answer to Question on Notice}\label{QON148}

\begin{figure}[h]
\includegraphics[width=0.8\textwidth, center]{QON148.png}
\end{figure}

\newpage
\subsection{Answer to Question on Notice}\label{QON150}

\begin{figure}[h]
\includegraphics[width=0.7\textwidth, center]{QON150.png}
\end{figure}

\newpage
\subsection{Answer to Question on Notice}\label{QON151}
\begin{figure}[h]
\includegraphics[width=0.7\textwidth, center]{QON151.png}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.7\textwidth, center]{QON151p2.png}
\end{figure}

\newpage
\subsection{Answer to Question on Notice}\label{QON1}
\begin{figure}[h]
\includegraphics[width=0.7\textwidth, center]{QON1.png}
\end{figure}

\newpage
\subsection{Answer to Question on Notice}\label{QON2}
\begin{figure}[h]
\includegraphics[width=0.7\textwidth, center]{QON2.png}
\end{figure}

\newpage
\subsection{Answer to Question on Notice}\label{QON3}
\begin{figure}[h]
\includegraphics[width=0.7\textwidth, center]{QON3.png}
\end{figure}

\newpage
\subsection{Hansard Extract, Senate Economic References Committee Public Hearing - 7th June 2018 \cite{SEN2018morrison}}\label{SENmorrison}

\textbf{Mr Morrison:} I would firstly like to thank and acknowledge two supporters: Mr Gary Johnston from Submarines for Australia, without whose support I wouldn't have been able to put together the substantial report that you have today; and my partner in Trendlock, the analytics consulting company that we used to deliver the report, whose help was invaluable.

In my opening remarks I would like to briefly restate the core conclusions only, and then I'd like to contrast and compare those with some other public statements that have been made on the record by Naval Group officials and officials of the Australian government. The key conclusions that I arrived at were that pump jets have a far lower efficiency than propellers at a low speed of travel in contrast to high speeds, where jets tend to become more efficient. The crossover does not lie near or anywhere under the probable transit speed of eight to 10 knots of a diesel-electric submarine. One world-renowned expert I discussed the matter with said it was impossible that a pump jet could be designed to be more efficient than a well-designed propeller under 18 knots.

One of the main reasons for the precipitous decline in propulsive efficiency at low speeds is directly linked to the mechanisms that enable the acoustic advantage to be achieved at high speeds. The duct has the effect of slowing the water down where the blades do their work on it. This has the effect of raising the pressure, which is what reduces the phenomenon of cavitation. Consequently, redesigning or modifying a pump jet to improve propulsive efficiency at low speeds cannot be done without necessitating some trade-off in terms of acoustic advantage achieved through the elimination of cavitation at higher speed. Very importantly, I also concluded that the extent of the efficiency declines at low speeds over the operational profile of a diesel-electric submarine will be substantial. An improvement in the order of 50 per cent or more in dive range or endurance could be achieved through switching back to conventional propellers. This could amount to some days dived on station or hundreds of nautical miles. In my report, I made the conclusive comparison that, between two otherwise identical submarines, the one with the pump jet will always have a lower dived endurance, a lower dived range, a worse indiscretion ratio, a lower overall endurance and a lower overall range than the one with the propeller. This statement, so far as I know, has yet to be challenged, and I'd be very surprised if anyone with technical competence or professional integrity sought to challenge it.

In contrast, I would like to quickly turn to a couple of other statements that have been made on the public record. The first is from the official Naval Group website, which says about the Shortfin Barracuda:

``Pump jet propulsion means the Shortfin Barracuda can move more quietly than submarines with obsolete propeller technology."

This is a very interesting statement because at face values it reads as if to say that pump jets are now so quiet that propellers have become obsolete. I know that that interpretation has been made, and I think it has been made before this committee or another Senate committee. Part of my further research that you have before you today indicated that the phenomenon that leads to increased turbulence at very low speeds is also likely to produce much more noise than a propeller would at very low speeds. This is a phenomenon called flow separation. It actually seems to me very surprising that the claim would be made that at very low speeds a pump jet would indeed be quieter than a well-designed propeller. Read more literally, it actually sounds like a different statement might be made, for the pump jet is only quieter when compared with an obsolete propeller. This could actually be the statement that was intended to be made, in which case it is perfectly consistent with the research that I have done and put before you today.

It might be very strange to try to figure out-and it was to me-why it would be useful to have a submarine that was only quieter at relatively high speeds and actually conferred disadvantages at all other speeds, including acoustically, with a significant trade-off in terms of overall range. It seemed like it would be quite a strange tactical circumstance that would require such a thing. In particular, and this is something else that needs emphasising, high speed for a submarine is actually still low speed for basically any other surface ship. Even very slow cargo ships moving at 18 or 20 knots move much faster than a submarine would be able to sustain for any significant period of time. Warships, of course, can easily go much, much faster, so it's difficult to see what that particular circumstance would be. I would like to emphasise that the committee should pay particular attention to what that circumstance might be.

Again, on the Naval Group public website, it is stated that:

``In a confrontation between two otherwise identical submarines, the one with pump jet propulsion always has the tactical advantage."

This is a very interesting statement as well. It certainly does not hold against a nuclear submarine. It wouldn't hold against any kind of surface ship. It doesn't hold in a strategic sense, either, because the necessity to snort more frequently on the way in and the lower dived endurance on station would mean that you'd be more exposed more quickly after any encounter. However, tactically, once the encounter has already happened and a target has been acquired, it is plausible that this statement is still true. It is quite clear that, since some considerable time ago, DCNS has been very, very well aware of what the advantages and disadvantages are of pump jet propulsion. There are also a number of other statements on the public record that also concur and fit extremely well with the research that I've done. In particular, statements by Jean-Michel Billig, who said back in October in 2017 that they may switch back to propellers and even insert AIP for higher underwater endurance, are perfectly consistent with the research that I've done. As well, more recently, the statements by Brent Clark on the ASPI Strategist blog published only a couple of weeks ago suggesting that a pump jet would not be suitable for a small submarine-but this submarine, being big enough, is perfectly consistent, as I pointed out in my public reply, with the research that I have already conducted. It seems to me that there is no contradiction clearly put by any representative of Naval Group to the research that I have presented on pump jets.

I would like to compare that briefly to a couple of statements that have been made by representatives of the Commonwealth. Perhaps the first that is worth turning to is that by Rear Admiral Greg Sammut:

``It can be efficient across the entire speed range, taking account of the submarine's size, the speed required and stealth. It needs to be tuned to the hull of the submarine. The hull needs to be tuned to the pump jet."

This was a statement made in October 2017, also published in an ASPI Strategist blog post his is a strange statement, because the first part-that it can be efficient across the entire speed range-taken on its own does not seem to be possible to be true. It seems to be completely false. Taking into account the submarine's size, the speed requirement and stealth, it seems to suggest that only if the admiral meant that he didn't actually want it to be efficient, and therefore could treat it as efficient-that seems to be the only way that it could be reconciled with reality and the physics that I've discovered.

The subsequent statements are even more peculiar - that it needs to be tuned to the hull of the submarine. As far as I can tell, the hull of the submarine proposed is so close in length and diameter to the one that's actually already being built by the French for the nuclear version, it seems very, very, strange indeed to make a tuning when such a small adjustment would be required. In contrast, if what was actually meant was that the pump jet should be tuned to the speed, the degree of the change in the speed of the operational profile would necessitate such a drastic overhaul to the design of the propulsion system that the idea of tuning seems quite thoroughly out of place. This is, to put it simply, as if someone bought an efficient hybrid car like a Prius and then decided to fit it with tank tracks from a dozer or a battle tank, and then argued they would tune the tank tracks back to be as efficient as the Prius was in the first place. It goes radically against what the technology was intended for in the first place. The idea of tuning is completely odd to any technical reading of the story.

I'd like to look at one further public statement that was made. That is from Stephen Johnson. This was made in relation to an answer that was given to this committee some time ago. When challenged very specifically about efficiency by Senator Patrick, he made this statement:

``No matter what kind of propeller or propulsion pump jet you design, it has to be optimised for a particular speed. There are classes of surface ships that look to optimise cross ocean transit. Physics usually keeps you around 17 to 18 knots, and then there is the cost of maintenance and the cost of repair and the fact that there's no particular stealth requirement. Typically, you will find that those ships will use an open-tip propeller. However, other ships, cruise ships in particular, use an electric shrouded pod, which is a variation of a pump-jet technology."

This also struck me as quite strange, because in my research I went to some lengths to make sure that I understood all the different variants of different types of propulsors, including propellers, ducted propellers, shrouds of various sorts and the way the different names are used. So far as I can tell, despite the fact that cruise ships typically travel substantially faster than would be possible for any diesel and electric submarine, this is particularly strange because I can't find any evidence of any propulsor used on a cruise ship that does indeed include a shroud. It seems to me that some confusion might have occurred about what a propulsor actually is in this case. To clarify, my best understanding is that 'propulsor' is a perfectly general term that can be applied to basically any basic pump-jet propulsor or any other device to push something through the water. It is sometimes used just to differentiate novel, different or contrasting systems.

I wonder whether here Mr Johnson has actually been confused with the Azipods that are installed on large cruise ships. They do use an electric motor, that's the only similarity I can find with the pump jet, and are indeed open to propellers and are quite similar to the propulsion systems used on the LHD. That is the best explanation that I can find so far for these comments. I certainly can't find any evidence of any shrouds being used on cruise-ship propulsion. Moreover, I can't find any evidence of pump jets of this sort being used on any ship, for which economic travel is a main priority, operating below the speeds of about 25 or 30 knots. So it's very interesting to wonder what exactly Mr Johnson was referring to in this case.

I will rest my opening statement there, but there are other statements as well and other matters to do with the way the procurement process and design process looks to be unfolding that seem concerning from the readings of the physics that I have done so far. I'm happy to have some further questions.

\textbf{CHAIR:} Thank you. I just want to place on record that since the commencement of the public hearing this morning Senator Reynolds has been joining us over the telephone line. She is now here with us.

\textbf{Senator PATRICK:} Mr Morrison, firstly I will point out that Mr Johnson is in the back of the room, so he'll have an opportunity later today to rebut anything that you've said.

\textbf{Mr Morrison:} Excellent.

\textbf{CHAIR:} We'll certainly give him that opportunity. In terms of your credentials-you haven't explained. Have you got engineering or scientific credentials?

\textbf{Mr Morrison:} I have a physics background. I studied theoretical physics at the ANU and got first-class honours in theoretical particle physics. I commenced a PhD at the University of Melbourne, which I didn't finish. I went to start a small defence technology company, designing small boats and torpedos for maritime warfare-a swarming concept. We used pump jets. We selected pump jets for the torpedo and water jets for the boat, so I became familiar with the particular systems and the trade-offs involved in those during that time. I had considerable discussions with manufacturers, naval architects et cetera during that process. I would like to point out that we selected pump jets for both those applications. I'm a fan of pump jets per se. They're very good. They're a go-to in many circumstances. It is a very peculiar, quite odd, situation to employ them on a conventional submarine.

\textbf{Senator PATRICK:} I just want to make sure that I've understood you correctly. You are saying that physics suggests that the pump jet is not efficient at the low speeds that a conventional submarine typically operates at in an operational area?

\textbf{Mr Morrison:} That's right.

\textbf{Senator PATRICK:} Of course, they use pump jets on nuclear powered submarines. Your paper says that's not a problem because they have no problems in respect of generating power, and you actually even say it's helpful.

\textbf{Mr Morrison:} Yes. Would you like me to clarify that point?

\textbf{Senator PATRICK:} Yes.

\textbf{Mr Morrison:} Nuclear reactors have a tendency to have different instabilities build up when they try to oscillate their power outputs radically. There are different decaying isotopes, xenon in particular, that make the reactor less stable after it has dramatically reduced its pressure. They prefer to operate in a steady state. They certainly wouldn't like to go down to very low power and then very high power and then very low power. That makes them somewhat more difficult to manage. So, to the extent of keeping a floor, a minimum level of energy required, they wouldn't need to pull the energy down below. That would be useful. They would reduce the extent that they would have to run other pumps or other systems to be able to take the excess energy out of the reactor.

\textbf{Senator PATRICK:} To your knowledge, are there any other conventional submarines-clearly, there are nuclear submarines that use a propulsor, or pump jets, but are you aware of any other conventional submarines that use them?

\textbf{Mr Morrison:} There are none in service at all. I'm aware of a prototype or demonstrator that was built once on a Russian submarine, a Kilo class submarine. I've seen a photo of that, but as far as I can tell it was never put into much active service.

\textbf{Senator PATRICK:} The bottom line for you is that it's problematic for a conventional submarine for all the reasons that you mentioned in terms of 'efficiency is king'. It affects indiscretion ratios and it exposes a submarine, because it will have to snort more often.

\textbf{Mr Morrison:} That's right. It seems like a genuinely strange choice. Once you take out the superpowers that a nuclear reactor confers on a submarine, leaving the pump jet is a massive hindrance to everything else about a diesel-electric submarine. The impact on the indiscretion ratio range, dive and endurance seems quite substantial.

\textbf{Senator PATRICK:} I tabled a document on 15 December, and it was pointed out that it was a 1963 declassified document from the US Navy. There were some criticisms levelled at me because of the age of the document. How reliable is the physics? Are you satisfied that the physics hasn't changed substantially, such that the information in the public domain could be wrong?

\textbf{Mr Morrison:} I'm absolutely satisfied. I have done quite a bit of research. As I mention in my blog post, I have referred to many more modern papers as well. The field of hydrodynamic propulsion does evolve; it does change. However, it tends to build on and confirm some of the most foundational and fundamental principles on which the conclusions that I am sure your paper and a few others that I have here with me, which are from, as you say, the sixties and seventies, are built.
It is particularly important when you are reading more modern literature that you find typically authors who wrote during those times who still consistently turn up in the references; they are still referred to. They are not dismissed in any regard. They are referred to as, I suppose, 'the greats' who wrote the books on these topics. They produced the full discussions of a comprehensive method for designing a pump jet. Even at that time, it was clear that pump jets never were intended - and I can give you some quotes, if you like, from certain old papers - for those very low speeds at all. In fact, they were really only advantageous at quite high speeds where cavitation was a severe problem.

\textbf{Senator PATRICK:} What you're saying is akin to saying that we shouldn't reject Einstein's theory of relativity just because it's old?

\textbf{Mr Morrison:} That's absolutely right. It's foundational, in fact. I can see nothing-and I have discussed this with other current experts-that has overturned the books, so to speak, on these matters. So I think some of those old foundational papers-and, indeed, I think I quoted one where it is surmised in a sentence-reached the conclusion: 'It's not good at low speeds'-almost as simply as that. They don't go back and re-labour and reprove what has been so long and well established in the annals of physics.

\textbf{Senator PATRICK:} You say that, acoustically, the pump jet is actually noisier at low speeds?

\textbf{Mr Morrison:} I think that's highly probable. I won't make the absolutely categorical claim, as I've done with the efficiency, but I think that is extremely highly probably because, as you can see from my efficiency research, a pump-jet would be producing something like twice as much energy release to turbulence at those very low speeds as what a propeller would be. It's hard to see how something discharging twice as much energy into the water can still remain quieter. Indeed, in this case, the sorts of turbulence are very large scale vortices. This phenomenon called flow separation and the periodic vortex shedding that might occur into the rotors, that's a relatively low-frequency sound; it's not the sort of high-frequency sound that might be associated with the very, very early onset of cavitation. For very high-frequency sounds, the shielding from the duct is likely to be effective-still only in the directions that are perpendicular to the access of the shaft, but will be somewhat effective. At the lower frequencies, I think there is good reason to think the shielding will not be as effective. I won't make the absolutely categorical claim that it has to be much noisier, but it seems difficult to believe how it is superior. I've never heard anyone explicitly make that claim, that it is actually acoustically superior at low speeds to a state-of-the-art propeller. I would be very surprised if that was the case.

\textbf{Senator PATRICK:} Chair, I would like to table a paper that was produced by DCNS in 2012 at a UDT conference? Will you permit me to do that?

\textbf{CHAIR:} Is there any objection to that document being tabled? No? That's okay.

\textbf{Senator PATRICK:} I would like the witness to have a look at it. I think the company was called DCN then, and it's their engineering area. Could you look on page 8 of this document? I've highlighted the relevant text. This is about the French Scorpene submarine and the permanent magnet motor. It says: 'The main advantages in terms of signatures of a permanent magnet motor is the ability to maintain a very low turns per knot, even with a high output, thanks to the high torques that are available for that particular motor. This low TPK reduces significantly the propeller cavitation apparition risks. For instance, within conventional operating depths, so 150 metres and higher'-which I presume is French for deeper-'no cavitation is expected whatever speed may be.' That tends to suggest to me that the French designed a propeller back in 2002 that does not cavitate at any speed provided you are below 150 metres. Is that consistent with your view on propellers versus pump-jets?

\textbf{Mr Morrison:} Yes, that's absolutely consistent. I think there's a diagram that I've drawn in my report, in the acoustics section. It's absolutely the case that raising the static pressure, which is what happens when you go very deep-and it's also what the shroud of the duct attempts to achieve to the water flow in a pump-jet-is what reduces the onset of cavitation, which is the principal cause of all the noise. Going deep simply does reduce or eliminate that. I think that, in this case, it is quite probable that at more modest speeds you'd be able to go quite shallow, potentially up to almost all of the routinely-used speeds of a diesel-electric submarine. The reason why you would want to be able to go in very shallow water at very high speeds for a very short period of time-it seems like a very narrow and a very obscure tactical advantage indeed, and that seems to be confirmed by the limitation of the superiority to a contest against another diesel-electric submarine only. It's hard to imagine how, against any other target, it could be useful.

\textbf{Senator PATRICK:} You've got an engineering background. You said you've worked on some projects. What do you see is the risk, if you were to categorise the risk of Defence achieving the publicly-stated performance expectations of the pump-jet-that is, efficiency across the entire speed range and acoustically quiet?

\textbf{Mr Morrison:} If they mean 'efficiency' to be anything close to a propeller across the entire speed range, I think achieving efficiency is impossible. I don't think they'll achieve that at all. I'd like to point out as well that there are a couple of other comments that seem particularly at odds regarding the engineering process and the design process that's happening here. It was stated by Rear Admiral Greg Sammut in the same post I referred to earlier. He said:
``We did not specify that the boat should have or not have air-independent propulsion or lithium-ion batteries, rather that it met our requirements. We'll continue to look at all applicable technologies as we go through the design process."

I think Brent Clark, in his very recent post on the ASPI Strategist, said the decision about the best propulsion for the submarine will be made by the Defence department. He is indicating quite clearly the final decision is yet to be made. And yet, still as of October, Admiral Sammut was saying:

``Work on establishing the dimensions of the submarine is complete. That pre-sizing-the initial length, diameter, displacement and other key parameters …"

It seems very clear that the size of the hull, the total displacement, the volume of the hull is now fixed. It seems very clear as well that the requirements are fixed, one would think.

However, that sits greatly at odds with the idea that the technology surrounding propulsion, particularly pump-jets, is not fixed. The reason for that is quite simple. As I've described through my research, the difference in efficiency is sufficiently large that to meet the same fixed requirement would differ the amount of fuel and batteries required in the order of hundreds of tonnes-hundreds of tonnes-for this sized submarine. It seems very difficult to imagine how, for a submarine, which basically has to be perfectly neutrally buoyant in all its operations, we could add or subtract hundreds of tonnes in the same fixed displacement without having a catastrophic engineering failure. I'm talking about Super Seasprite style: build the air frame, set the requirements. They simply do not go together whichever way you put them. The risk that you simply have a submarine that doesn't float or doesn't sink-I'm not sure which is worse-is abundantly clear from these. The only alternative that I can see is that they actually haven't fixed the requirements. They've designed the hull and they'll have whatever performance turns out of whatever technologies they put into it in whatever volumes seem suitable.

\textbf{Senator PATRICK:} I think you've partly answered this, but what do you see the risk is for the submarine project of proceeding with the pump-jet, noting you say it's impossible to achieve everything that they want to achieve and particularly noting the weight, the size and the fact that it's positioned on the arced extremity of the submarine?

\textbf{Mr Morrison:} I think it adds exactly the point I've made: This is a very, very major change. I didn't even mention that; it's in addition to the additional total volume of fuel and batteries required. Locating something that large and heavy, the shroud will be many tonnes, tens of tonnes, at the rear end. Quite frankly, it looks incomprehensible. I don't know how you could consider saying that you've just won a concept about a submarine if you haven't defined the propulsion system or the type of battery or whether it has got AIP. There is simply nothing more important in the engineering of the whole thing than fixing those sorts of things in place. It seems extremely odd to me that they are saying they're still considering these things and they've fixed the size of the submarine and they're making progress on the design, because all these things will have such a radical impact on the amount of weight embarked and the balance of the whole thing. It seems like an incomprehensible sequence of events to have happening at this stage.

\textbf{Senator PATRICK:} Thank you. That's all I have.

\textbf{Senator KIM CARR:} Mr Morrison, it's a very important point you make. You're going to the very fundamental design questions that are at the heart of a major acquisition for this country. I'm just wondering what are you seeking from this committee in terms of the propositions you're putting before us?

\textbf{Mr Morrison:} Look, I'm just offering my evidence here. To me it seems that the pump-jet is strange. I find it difficult to believe that this is the primary design feature that we have knowingly and deliberately selected. Whatever has led to that, I just suggest this committee needs to get to the bottom of why it was selected because it doesn't seem consistent with building a world-class submarine and it seems like we could be heading on a track that could lead to a real procurement mess. I would just direct your focus of figuring out how we got here, because if the forces that put such a strange engineering choice into the submarine are still at work then there could be worse to come in the future and it doesn't bode well for a successful procurement.

\textbf{Senator KIM CARR:} Given where we are in the procurement process, I think it is incumbent upon us to determine what we do with your evidence. What are you actually recommending we do? For instance, if I might be so bold: I'm not certain this committee would have the technical expertise to evaluate what you're saying, and the Department of Defence clearly challenges what you're saying, so where do we go to evaluate what you're proposing?

\textbf{Mr Morrison:} So far as I can tell, this is not controversial. Whilst it seems to be very technical physics, the paper that you -

\textbf{Senator KIM CARR:} I'm not certain the Department of Defence would agree with your proposition that it's not controversial.

\textbf{Mr Morrison:} I'd be very interested to see that. I've had discussions with other concerned physicists and people in the naval community about this, and I've heard of discussions had with the Department of Defence and some of the basis for their thinking that pump jets are as efficient, plausibly, over the whole speed range. I'm quite confident that further investigation needs to be done of that. I've been pointed to a particular chart, and I think they've misread it; I think they haven't been able to interpret it. So, I think you should dig hard. I'm quite confident of the research. I think that my paper to you is a carefully and meticulously researched statement of what is relatively obvious. I'm not aware of anyone in the marine engineering community seriously proposing that you would ever select a pump jet for something designed to operate anywhere near, let alone below, 10 knots. It is quite a strange proposition, so I think you should investigate it very, very seriously indeed.

\textbf{Senator KIM CARR:} You're seeking from this committee further investigation of your claims. Is that the essence of your recommendation to us?

\textbf{Mr Morrison:} Yes, and I would also strongly suggest that you carefully investigate the basis for the anti-submarine-warfare superiority claim. To me, that seems the last plausible thread of logic on which you could justify a submarine choice with a pump jet, and that's to say last plausible and not categorically and obviously untrue. There are gross doubts that I have about that as well: how you could ever acquire such a small target passively; how you would measure the range that you were at and know that you would be able to close in the short distance that you're able to; whether you'd be able to keep track of your target when you accelerate to those high speeds-there are a whole bunch of reasons which look almost as dubious as the proposition about pump jets being efficient.

I'm actually seeking to do some more research on that topic exactly, because it seems perfectly clear that that is the only situation in which there's a plausible case that the pump jet might be of any value, and I think your focus should very much dwell there. I'd like to get some more resources to conduct a thorough study of that, and I think that it should be challenged very directly with Defence-whether we're buying a submarine for all our strategic needs or whether we have just selected one particular scenario to justify a choice that's based on some other justification. I can't see a logical justification for having made this particular design choice other than that. That is the last plausible reason there might be, and it's only barely plausible. I think it should be investigated very thoroughly.

\textbf{Senator KIM CARR:} You're saying you're undertaking further research. What's the state of that research?

\textbf{Mr Morrison:} It's very early. I'm basically seeking further financial support to be able to do the same sort of sustained and thorough investigation as in the paper that you've got before you today. I haven't been able to get that yet, but I think it would be worthwhile. I think it would be in the interests of the Commonwealth to determine whether there is serious merit in that statement. I wouldn't make the sort of categorical and strong claims I've made about efficiency about the anti-submarine scenario without doing that sort of research. I like to take the academia, the literature and the physics very seriously. But I think that sort of thing should be done quite urgently because I think there have been some very strange, potentially misleading-perhaps deliberately misleading-statements made about the superiority of the pump jet.

\textbf{Senator KIM CARR:} This committee is due to report on 27 June. I'm not certain we'd have a lot more time to -

\textbf{Senator PATRICK:} Senator Carr, if I may, just to kick this along a bit-it's a very good thought you're having. Do you have a security clearance?

\textbf{Mr Morrison:} No.

\textbf{Senator PATRICK:} That would preclude you from getting in. My thought was you could perhaps go inside, work with Defence and then pop out and say either, 'I'm not convinced,' or, 'I am convinced,' keeping it at an unclassified level.

\textbf{Mr Morrison:} The rules of physics are not classified or classifiable, and I think that's -

\textbf{Senator KIM CARR:} I'm interested in pursuing what you're saying. As a citizen, I think you're entitled to put views which go to the heart of this procurement. I'm not saying that you're right. Please don't misinterpret me-I'm not qualified to say that you're correct, but I am interested in what you're saying because of the seriousness of the propositions you're advancing. I am trying to come to grips with what we do with what you're saying. We've only got limited time. I'm clearly of the view we're about to face a general election, and it may well be that there will be a different set of circumstances after that election. In terms of this immediate report that's required to go to the parliament, we need to find a way to evaluate what you're saying, given that this committee, in my view, doesn't internally have the capacity to technically evaluate your claims.

\textbf{Mr Morrison:} Might I suggest, as a way to trying to seek validation, you could see whether anyone else would contradict or oppose some of the simple statements that I've made.

\textbf{Senator KIM CARR:} We'll ask the Department of Defence but it will not surprise me if they tell me there's a technical reason why you can't be right.

\textbf{Mr Morrison:} I'd be amazed to see that. Having been challenged, I would be amazed if someone wants to restate-and I've made some careful statements in my report. You can ask them whether I'm wrong and see what they come back with. I'd need to be quite badly wrong on most of this.

\textbf{Senator KIM CARR:} Would you accept it on technical grounds, if they can demonstrate that you're wrong in theory?

\textbf{Mr Morrison:} The whole thing has got to be wrong. It's got to be utterly wrong. I think this is worthwhile. If I was to go to a particular page of my report-on page 32, if you like, there's a chart from a paper that was done not too many years ago by BMT Defence Services-you can see the graph there-that shows the efficiency curve of a fully submerged water jet and a propeller going down to low speeds. Just look at the shape of the curve and the gap between those. If the department comes back and says I'm wrong on some technicality, some small thing, which means the lower curve is not exactly right and it needs to be shifted up a tiny bit, that would simply not matter at all. The conclusions I am putting to you would still be right. They need to argue that the curve literally does not decline anything like that point, and I don't think anyone would try to make that case.

\textbf{Senator KIM CARR:} I think that's a reasonable proposition. We may well seek from the department a critique of your paper -

\textbf{Mr Morrison:} I'd be interested if they say it's not exactly the right design.

\textbf{Senator KIM CARR:} in the normal peer review process.

\textbf{Mr Morrison:} I am very happy to submit to my peers on this matter. I'm really quite confident. Or you might consider that this justifies further inquiry or extending this inquiry or a Senate inquiry.

\textbf{Senator KIM CARR:} That's right. But it may well be that the department agrees that it requires further inquiry. Is that what you're suggesting to us?

\textbf{Mr Morrison:} I think that when they check their numbers, yes, they might agree.

\textbf{Senator KIM CARR:} I look forward to that. Thank you very much.

\textbf{CHAIR:} Thank you very much, Mr Morrison.

...

\textbf{CHAIR:} I now welcome Mr Brent Clark from Naval Group Australia. Information on parliamentary privilege and the protection of witnesses and evidence has been provided to you. I invite you to make a brief opening statement and then we'll open it up for questions.

\textbf{Mr Clark:} Good morning, everybody. I'm the special advisor on the Future Submarine program and to the Naval Group Australia chairman, Jean-Michel Billig. I was formally the interim chief executive of Naval Group Australia. It is always a pleasure to appear before this committee to provide you with an update on the Future Submarine program from the perspective of Naval Group Australia. Naval Group's work on the program has been proceeding at pace since my last appearance here almost 12 months ago. What I intend to do today is bring you up to speed with our efforts to deliver Australia a regionally superior submarine that will be built, operated and sustained with Australian sovereignty.

Australia's \$1.6 trillion economy is dependent on secure sea lanes. As such, Australia needs the best possible submarine to protect our trade and support our maritime security. Working in collaboration with the Commonwealth and Lockheed Martin Australia, that is exactly what Naval Group will deliver. The Future Submarine program is the largest defence program in Australia's history and a truly national endeavour that will maximise Australian industry involvement and create thousands of jobs for many generations of Australians for decades to come.

Naval Group, in partnership with Combat System Integration, Lockheed Martin Australia and the Australian government, is building 12 regionally superior diesel powered submarines for the Royal Australian Navy in order to maintain Australia's regional maritime superiority and sovereignty. Submarines are one of the most complex artefacts on the planet. Each submarine will be made up of at least one million parts, compared to a car which has 3,000 parts. As a result, we need to take our time. This program cannot be rushed. We need to get it right. I can tell you today that Naval Group is more than halfway through the design and mobilisation contract which focuses on early design activities in coordination with Lockheed Martin Australia and our other Australian partners.

The Future Submarine is currently progressing through concept design on schedule. This has involved successful work to establish the initial length, diameter and displacement of the submarine ahead of the subsequent design phases. We are working with the Commonwealth on the components and systems that will be required for the submarine. The first stage of the submarine design, known as the functional analysis, was completed on schedule in October 2017, culminating in a successful preliminary system requirements review. We have finalised our Australian Industry Capability Plan, which will detail our early work to honour our commitments to maximise Australian industry involvement, particularly including SMEs.

As at the start of May 2018, 813 companies had engaged with Naval Group nationally through the Industry Capability Network, an online database which gives companies access to projects large and small. Suppliers for the top five pieces of equipment, including the main motor, diesel generators, switchboards, batteries and weapons discharge systems are currently being reviewed. We will work with the suppliers of this equipment to ensure that there is a full technology transfer and that Australian industry is able to maintain and support this equipment.

To give you some numbers to work with, the Naval Group workforce has already grown to about 90 people at our Adelaide headquarters, we have six engineers currently in France and we'll be recruiting another 31 engineers to transfer to France this year. That number will grow to about 100 Australian engineers in France over the next few years. Our workforce is expected to grow to about 155 by the end of this year. Also by the end of 2018, we're working with the Commonwealth towards turning the first sod of soil and laying the foundations for the new submarine construction yard at Osborne in Adelaide's northern precinct. Over the next three years, the construction sheds, which will house the equipment to build the submarines, and the construction halls for the submarines will go up. The establishment of the infrastructure in this time frame will be required to ensure that construction of the submarines can begin from 2022-23. We envisage building up the Naval Group Australia workforce to around 1,500 employees by around 2028-29. Overall, the program is expected to generate an annual average of around 2,800 jobs in Australia.

The first submarine will commence service in the early 2030s, with the last to be constructed some time in the late 2040s and sustainment to continue to the end of the 2080s. Over the next four years, our partners, Lockheed Martin, will establish the combat system testing and integration facility to test the capabilities that will contribute to the submarines' lethality and regional superiority. Naval Group will build a land based propulsion and testing facility to derisk the submarine's propulsion system, which is a sophisticated piece of engineering that attempts to reduce the noise signature of the diesel-electric propulsion system.

At this point, if the chair is happy, I'll table the remainder of my speech so that we can get to the questions.

\textbf{CHAIR:} Thank you very much, Mr Clark.

\textbf{Senator PATRICK:} Firstly, it's probably proper that we give you an opportunity to respond to Mr Morrison's evidence. That's probably proper, Chair.

\textbf{Mr Clark:} Thank you, Senator Patrick. I believe this is the first time we've sat opposite each other in a committee, so let the fun begin! From my perspective, I don't intend to offer any commentary on the previous speaker. Suffice to say that the information that we've been supplying of a classified nature to the Australian government is held by the Australian government and was supplied by the French government. The Australian government and the Australian Department of Defence obviously have access to classified research from the United States Navy, the Royal Navy and the French Navy on this matter, and I don't intend to get into a debate on this topic.

\textbf{Senator PATRICK:} Thank you. I might just step back. Mr Morrison suggested that there had been a Russian commercial submarine that had been fitted with a pump-jet.

\textbf{Mr Clark:} Yes.

\textbf{Senator PATRICK:} Can you confirm that the French fitted, for a trial, a pump-jet on to an Agosta 70 class submarine?

\textbf{Mr Clark:} Yes they did, many years ago.

\textbf{Senator PATRICK:} DCNS obviously has nuclear customers, being Brazil and, of course, the French Navy, and a range of conventional customers. Has DCNS offered or put a propulsor on any conventional submarine?

\textbf{Mr Clark:} Can I clarify one thing? Naval Group does not have a nuclear customer in Brazil. We are assisting the Brazilians to create their own sovereignty in their own country, but France has not committed to supply any kind of nuclear technology to the Brazilians.

\textbf{Senator PATRICK:} So they're doing the reactor-

\textbf{Mr Clark:} Correct.

\textbf{Senator PATRICK:} but you're doing the submarine?

\textbf{Mr Clark:} We're assisting them in terms of designing their submarine.

\textbf{Senator PATRICK:} Are you assisting them with a pump-jet?

\textbf{Mr Clark:} Not at this stage. Pump-jet technology, like much of the technology on this submarine, is incredibly sensitive. The DGA, the French authorities, are the design authority for the pump-jet for Naval Group and they are working with the Commonwealth of Australia. This is not a technology that's offered to anyone. This is a technology that is only offered to certain allies.

...

\textbf{Senator PATRICK:} Sure. I have one more question, and this is to give you an opportunity to respond to Mr Morrison's evidence, Mr Johnson. He contradicted your evidence about cruise ships having pump jets. He said that you might have meant Azipods?

\textbf{Mr Johnson:} I don't remember that clearly enough. I do remember that you had produced a document, and I was trying to answer questions without the benefit of where we were going.

\textbf{Senator PATRICK:} Just giving you an opportunity to-

\textbf{Mr Johnson:} I don't have a clear idea of exactly the choice of words I used, but I think the concept remains the same: propellers and propulsers are designed for very, very specific purposes, and if you change the purpose you will change your choice of design.

\textbf{Senator PATRICK:} Thank you.

\textbf{Senator KIM CARR:} Mr Johnson, you sat through the evidence this morning. I noted you at the back of the room when Mr Morrison made his presentation-is that correct?

\textbf{Mr Johnson:} Yes, I was here this morning. I wasn't necessarily paying attention.

\textbf{Senator KIM CARR:} You weren't? That's a big thing to admit on the Hansard!

Senator Reynolds interjecting-

Senator Kim Carr interjecting-

\textbf{Mr Gillis:} I will support my colleague: not all of this is riveting!

\textbf{Senator KIM CARR:} Is that right?

\textbf{Senator REYNOLDS:} Mr Gillis!

\textbf{Senator KIM CARR:} Mr Chesworth, maybe this is a point where it will be easier than you thought, with this level of humour! What's your response to the suggestion that we get a critique from the department as to the paper that's been presented to this committee? Is that possible?

\textbf{Mr Johnson:} I'm sure it's possible, yes.

\textbf{Senator KIM CARR:} We have a deadline in terms of the committee's reporting date. If you were conscience of that, would it-

\textbf{Mr Johnson:} What would that deadline be?

\textbf{CHAIR:} 27 June.

\textbf{Mr Johnson:} Surely you would want something before that?

\textbf{Senator KIM CARR:} Of course. That's the point. I'm not as humorous as you fellows, but it would look pretty bloody stupid after the report's handed in, wouldn't it?

\textbf{Mr Johnson:} I've only had a chance to look through it. First of all, we welcome all ideas. I think I don't want to say much more than we'll do our very best to support you.

\textbf{Senator KIM CARR:} The reason I raise this is that I think citizens are entitled to put a view about matters of this importance.

\textbf{Mr Gillis:} Absolutely.

\textbf{Senator KIM CARR:} I contend that we don't have the technical expertise to evaluate the report. It may well be that we recommend that further work be done on it, but I'd want to see your response to this-to give us a critique of it; your perspective as to the report. He's made some assertions to the committee. You no doubt will get the Hansard pretty quickly, if the secretary can make sure that that's available to the officers.

\textbf{Mr Gillis:} The only caveat I'll put on any response we make is that the one thing we do really well is detailed engineering analysis. When you have somebody's report-and I have not read the report yet; we weren't actually given a copy. To actually go through it-and it looks like something which is yea thick-with serious analysis is going to take us time. We can only do what we can do.

\textbf{Senator KIM CARR:} If anyone comes to me and says that fundamentally you've got it wrong on a technical matter such as a propulsion system, (1) we have to take notice of that, but (2) I'd be surprised, frankly, given the level of technical expertise available.

\textbf{Mr Gillis:} Senator, that's one. I regularly get correspondence from the public saying that almost every decision that we've made for the last 50 years on a defence matter is wrong, because somebody has an opinion.

\textbf{Senator KIM CARR:} Not since the fall of Singapore, surely!

\textbf{Mr Gillis:} Almost since the fall of Singapore. There are many people out there with opinions, and the vast majority-I would say over 99 per cent of them-you can dismiss within the first page of reading them because they are fundamentally wrong.

\textbf{Senator KIM CARR:} We clearly don't do that, because Mr Morrison wouldn't have got a hearing at this committee if we'd dismissed it. I'd like an assessment from you.

\textbf{Mr Gillis:} I accept that. We will provide the committee with the best possible advice we can in the shortest possible time frame to meet your requirements so that you can be better informed, because we do hold that expertise.

\textbf{Senator KIM CARR:} Thank you. There was a reference to advice that you have from various navies-obviously, classified advice. Now, you won't be able to draw on that, but we need to know whether or not it's consistent with the advice that you have in terms of contemporary-that's the point.

\textbf{Mr Johnson:} The intended plan is to use a propulsor. Is that what you're referring to?

\textbf{Senator KIM CARR:} The point of the solution-

\textbf{Mr Johnson:} That's well supported across-

\textbf{Senator KIM CARR:} It may be that we seek an in camera, private briefing, from you.

\textbf{Mr Johnson:} That's exactly where I was headed. That use is-I think it is fair to say-well supported. I don't want you to misunderstand me, because the very first thing I said was, 'Well, of course, we will.' But this report also talks about things-and I'll give you one example-like batteries and battery efficiencies. Those are not things we want out in the public-dive and endurance and those sorts of things. So the beauty of a paper such as this, which comes from outside of the establishment, is that there are sometimes good ideas in new papers. The disadvantage is that sometimes they are created and constructed with a misunderstanding-

\textbf{Senator KIM CARR:} Or without contemporary knowledge.

\textbf{Mr Johnson:} Without contemporary knowledge or without understanding necessarily the intricacy. For example, the shaft horsepower-the power required to move even a very big ship at very low speeds-is astonishingly small. If you look at inefficiencies, you look at a curve and say, 'That must be a problem.' It's not. It's like: 'Hey, you paid twice what you needed to for that. It was a penny. It cost you two pennies. That's 100 per cent over.' So some of this is out of context with what we're doing, which will make it hard to give you an answer of empirical-

\textbf{Mr Gillis:} One of the things that we learned during the competitive evaluation process and through all of the studies I've done on submarines-and I have nowhere near the life experience that the admiral here has-is that every country in the world is highly sensitive about the performance of their submarines, what technology they use and how they do it.

\textbf{Senator KIM CARR:} So they should.

\textbf{Mr Gillis:} So it is almost that the vast majority of external papers are written in a vacuum of information about what's really out there. That's the problem. The same thing happened on the joint strike fighter with stealth technology. We had 10 years of commentators saying the thing would never fly. It would never meet stealth requirements. It would never do those things. The only thing I haven't heard in the three years since I've been here was after the JSF landed here in Australia: they seem to have shut up, because it exists. So I just counsel that sometimes, especially on very highly sensitive defence materiel, people will use 'cousin Google' to get a lot of information. I'm not referring to this individual, but I see a lot of that stuff.

\textbf{Senator KIM CARR:} I don't think this fellow meets that criteria.

\textbf{Mr Gillis:} No, I don't. We have to value the input that we get from the public, and I look forward to reading the paper.

\textbf{CHAIR:} Our time is nearly out, so I will call on Senator Reynolds.

\subsection{Australian Defence Magazine, August 2018 Extract}

An extract from "Future Submarine: Pumpjet propulsor technology choice still not well explained" \cite{kerr2018}, article commencing Page 20.

As explained by Dr Kershaw, there was a degree of misunderstanding concerning pumpjet design and performance.  To propel a surface ship or submarine involved overcoming resistance to push the vessel through the water with a suitably-designed propulsor.
Surface ships need to overcome wave-making resistance, which varies with speed.  Their propulsors therefore need to be optimised for the required operating speed; when operating at off-design speeds, efficiency drops off.
Since submerged submarines do not have to overcome wave resistance, the efficiency  of a propulsor can be considered independent of speed and the aim is to design an efficient propulsor.
In other words, selecting the propulsor for a submarine is driven by the types of mission it is expected to undertake and where the boat will operate, Dr Kershaw stated.

``The question of propeller versus pump jet really comes down to the acoustic signature; a well-designed pump jet is considerably quieter than an equivalent propeller." he said. "One of the advantages of a well-designed pump jet is that it can smooth out the water flow that is coming off the back of a submarine before the flow hits the moving rotor and there's less unsteady flow, and thus less noise.
``This smoothing-out also increases the propulsive efficiency; a well-designed pump-jet can have efficiencies equal to, or better than, an equivalent propeller.  You just have to have the capabilities and facilities to be able to design an efficient pumpjet and we know that France can do that."
Echoing RADM Sammut, Dr Kershaw stressed that the pumpjet on a submarine must be designed in conjunction with the hullform to which it was to be attached.
``Its' not like a propeller where you can design your propeller open loop i.e. without considering the interaction with the hull, and then attache the propeller to the hull.  The water flow over the hull to the pump jet is so integral you've got to design it as a pair, and the hull and the pump jet have to be tuned together, it's a total system that you're building."

``One of the issues with pumpjets and the reason they're not used on diesel-electric submarines is they're much heavier than a conventional propeller, although they have a smaller diameter.
``As the size of the submarine increases (the Future Submarine's 4,700 tonne submerged displacement is significantly heavier than the 3,407 tonnes of the Collins-class it will replace) you start to get to the point where a pump jet can become sensible."

Although pump jet propulsion would be new to the RAN, the technology was well-known to DSTG.
"We've invested heavily in our hydrodynamics and hydroacoustics program over the past 20 years and this is includes the computational models as well as experimental facilities that confirm modelling results," Dr Kershaw said to \textit{ADM}.
We've been able to develop our analysis on a generic pumpjet design and thus confirm our understanding of pumpjet performance; you basically build it up from first principles.
"Our analysis at the moment is all around analysing current options but we will be looking at what's ahead in the future and enable Australia to develop the sovereign capability to take best advantage of pumpjet technology."

\textbf{Classified information}

It was worth noting, however, that modern pumpjets were primarily employed on military submarines and torpedoes and information was very limited and tightly held, making it difficult to review performance predictions.
``DST though, is in the position where we can access certain privileged information, so we can get validation of our analysis capabilities," Dr Kershaw noted.
Pumpjet design had advanced over many years with a particular forcus on the characteristics of all the components, although much of that was classified.
Some propellers were highly specialised and highly complex. Pump jets had complexities in other aspects but the moving rotor was not as complex as a normal propeller, benefitting sustainment.

So will pump jet technology definitely be used for the Future Submarine?
``That is what is currently in the design  and that's what a lot of our science and technology program is about; having the capability to do the independent analysis on the technologyies and capabilities planned for the Future Submarine, including the pump jet program," Dr Kershaw said to \textit{ADM}.
Discussions were currently underway with Naval Group and the French government regarding closer collaboration on a range of technologies and this would include pump jets, Dr Kershaw disclosed.
The program was still being worked on, and on the Australian side would involve a government-only team to comply with French security regulations.
``One thing I would like to emphasis is that Australia's geographic circumstances require a larger conventional submarine than those currently produced for export to other parts of the world.  The size of the Future Submarine means a pumpjet is a real game changer for optimising the performance of the submarine in its intended role, and ensuring regional superiority in undersea warfare."
Dr Kershaw concluded.

\newpage
\subsection{Defence Submission to Senate Economic References Committee}\label{def2018}

\begin{figure}[h]
\includegraphics[width=0.9\textwidth, center]{def1.png}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.9\textwidth, center]{def2.png}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.9\textwidth, center]{def3.png}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.9\textwidth, center]{def4.png}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.9\textwidth, center]{def5.png}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.9\textwidth, center]{def6.png}
\end{figure}

\end{appendices}

\end{document}